### 2016-12-15 paper/PacketSamplingFirewalls.tex
Answered a bunch of questions in notes/, started moving paper and proposal to LaTeX

\documentclass[fleqn]{article}%
\usepackage{amsmath}%
\usepackage{amsfonts}%
\usepackage{amssymb}%
\usepackage{graphicx}
%-------------------------------------------
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{algorithm}[theorem]{Algorithm}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\newenvironment{proof}[1][Proof]{\textbf{#1.} }{\ \rule{0.5em}{0.5em}}
\setlength{\textwidth}{7.0in}
\setlength{\oddsidemargin}{-0.35in}
\setlength{\topmargin}{-0.75in}
\setlength{\textheight}{9.2in}
\setlength{\parindent}{0.0in}
\begin{document}

\begin{center}
  \textbf{Packet Sampling Firewalls \\ Austin Voecks}\\
\end{center}

\textbf{Outline}
\begin{itemize}
    \item Problem
    \item Abstract
    \item Subsampling
    \item Randomized Subsampling
    \item Dynamic Subsampling
    \item Heuristic Based Rules
    \item IPFW
    \item Benchmarking Performance
    \item Results
    \item Security Implications
    \item Related Work
    \item Conclusion
    \item References
\end{itemize}

\textbf{Problem} \\
Provide non-disruptive backend network monitoring to OneFS at 40Gb speeds.

\textbf{Abstract} \\
The performance of FreeBSD’s IPFW can be increased to support 40Gb network
interfaces without additional specialized hardware through dynamic
probabilistic rule matching. The probability of rule matching will be based on
current system load, allowing it to adapt to changing demands on the system.
The efficacy of packet stream subsampling will also be explored.

\textbf{Subsampling} \\
Ideally, all network traffic would be carefully inspected; this is impractical
due to the limitations of CPU and bus speeds. When logging, it’s also limited
by disk speed. Subsampling traffic provides a way take the temperature of the
network and increase chances of detecting abnormal behavior; some monitoring is
better than none. To provide full packet inspection to every packet on system,
the network links would have to be split into 10Gb links, each passed through
dedicated hardware.

\end{document}

### 2016-12-15 paper/main.tex
Answered a bunch of questions in notes/, started moving paper and proposal to LaTeX

%% bare_jrnl.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{IEEE Journal Template Example}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Michael~Shell,~\IEEEmembership{Member,~IEEE,}
        John~Doe,~\IEEEmembership{Fellow,~OSA,}
        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
\thanks{M. Shell is with the Department
of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
\blindtext[1]
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
IEEEtran, journal, \LaTeX, paper, template.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
\blindtext

\subsection{Subsection Heading Here}
\blindtext

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

\section{Conclusion}
\blindtext

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
\section{Proof of the First Zonklar Equation}
Some text for the appendix.

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2016-12-15 paper/compile.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+pdflatex -output-directory=/tmp/ PacketSamplingFirewalls.tex && cp /tmp/PacketSamplingFirewalls.pdf ~/Downloads

### 2016-12-16 ipfwd/ipfwd.c
Update comments, readme for ipfwd/

@@ -1,12 +1,16 @@
+/* ipfwd.c 
+ * author: Austin Voecks
+ *
+ * Monitors system load using uptime() to set probability of IPFW's first rule
+ * allowing any packet. This rule increases performance by sacrificing complete
+ * security. The probability is dynamic based on the current load.
+ */
+
 #include "ipfwd.h"
 
 double sigmoid(double x) {
   /* push low values lower and high values higher
-
-     sigmoid(0.1) = 0.039
-     sigmoid(0.5) = 0.500
-     sigmoid(0.9) = 0.961
-     */
+   */
   if (x < 0.1) return 0;
   if (x > 0.9) return 1;
   return 1 / (1 + exp((x - 0.5) * -8));
@@ -14,7 +18,7 @@ double sigmoid(double x) {
 
 void control_ipfw(enum command cmd) {
   /* enables or disables ipfw
-  */
+   */
   int enabled, on = 1, off = 0;
   size_t size = sizeof(enabled);
 
@@ -41,7 +45,7 @@ void control_ipfw(enum command cmd) {
 
 void set_probability(double probability) {
   /* run the ipfw commands to set the new first rule probability
-     */
+   */
   int status;
   pid_t pid;
   char *rule_body = "-q add 1 prob 0.000 allow ip from any to any ";

### 2016-12-16 ipfwd/ipfwd.h

@@ -1,3 +1,11 @@
+/* ipfwd.h
+ * author: Austin Voecks
+ *
+ * Monitors system load using uptime() to set probability of IPFW's first rule
+ * allowing any packet. This rule increases performance by sacrificing complete
+ * security. The probability is dynamic based on the current load.
+ */
+
 #include <stdio.h>
 #include <stdlib.h>
 #include <unistd.h>

### 2016-12-16 loadgen/loadgen.c
Update comments for loadgen/

@@ -1,3 +1,10 @@
+/* loadgen.c 
+ * author: Austin Voecks
+ *
+ * Program that simluates CPU load by doing some math and then waiting for an
+ * empirically determined number of nanoseconds. 
+ */
+
 #include <math.h>
 #include <time.h>
 #include <stdio.h>

### 2016-12-16 downloads/CPULoadGenerator/CPULoadGenerator.py
Remove CPULoadGenerator

@@ -1,65 +1,0 @@
-#!/usr/local/bin/python
-
-#Authors: Gaetano Carlucci
-#         Giuseppe Cofano
-
-import multiprocessing
-from twisted.python import usage
-
-import sys
-sys.path.insert(0, 'utils')
-
-from Monitor import MonitorThread
-from Controller import ControllerThread
-from closedLoopActuator import closedLoopActuator
-
-class Options(usage.Options):
-    """
-       Defines the default input parameters
-    """
-    optParameters = [
-            ["cpuLoad", "l", 0.2, "Cpu Target Load", float],
-            ["duration", "d", 10, "Duration", int],
-            ["plot", "p" , 0, "Enable Plot", int],
-            ["cpu_core", "c" , 0, "Select the CPU on which generate the load", int]
-        ]
-                 
-if __name__ == "__main__":
-
-    import sys
-    options = Options()
-    try:
-        options.parseOptions()
-    except Exception, e:
-        print '%s: %s' % (sys.argv[0], e)
-        print '%s: Try --help for usage details.' % (sys.argv[0])
-        sys.exit(1)
-    else:
-        if options['cpuLoad'] < 0 or options['cpuLoad'] > 1: 
-            print "CPU target load out of the range [0,1]"
-            sys.exit(1)
-        if options['duration'] < 0: 
-            print "Invalid duration"
-            sys.exit(1)
-        if options['plot'] != 0 and options['plot'] != 1: 
-            print "plot can be enabled 1 or disabled 0"
-            sys.exit(1)
-        if options['cpu_core'] >= multiprocessing.cpu_count(): 
-            print "You have only %d cores on your machine" % (multiprocessing.cpu_count())
-            sys.exit(1)
-    
-    monitor = MonitorThread(options['cpu_core'], 0.1)
-    monitor.start()
-
-    control = ControllerThread(0.1)
-    control.start()
-    control.setCpuTarget(options['cpuLoad'])
-
-    actuator = closedLoopActuator(control, monitor, options['duration'], options['cpu_core'], options['cpuLoad'], options['plot'])
-    actuator.run()
-    actuator.close()
-
-    monitor.running = 0;
-    control.running = 0;
-    monitor.join()
-    control.join()

### 2016-12-16 downloads/CPULoadGenerator/tests/Identification.py

@@ -1,89 +1,0 @@
-#!/usr/bin/python
-
-#Authors: Gaetano Carlucci
-#         Giuseppe Cofano
-
-
-import json
-import matplotlib.pyplot as plt
-
-import sys
-import os
-sys.path.insert(0, os.path.dirname(os.path.abspath(__file__)) + '/../utils')
-
-from Monitor import MonitorThread
-from openLoopActuator import openLoopActuator
-
-period = 0.05 
-
-def cpu_model(cpu_target):
-    cpu_time = cpu_target * period
-    sleepTime = period - cpu_time
-    return sleepTime
- 
-if __name__ == "__main__":
-    cpuSequence = [0.1, 0.8, 0.30, 0.70, 0.40, 0.10, 0.20, 0.60, 0.20, 0.70]
-    ######################################################
-    #             IDENTIFICATION TEST                    #
-    ######################################################
-    # testing activities
-    # this test aims at characterizing the CPU
-    testing = 1
-    dynamics_plot_online = 0
-    if testing == 1:
-    
-        sleepTimeTest = [ cpu_model(x) for x in cpuSequence]
-
-        data = {"x":[], "y":[]}
-        for sleepTime in sleepTimeTest:
-            monitor = MonitorThread(0, 0.1)
-            monitor.setSleepTimeTarget(sleepTime)
-            monitor.start()
-
-            actuator = openLoopActuator(monitor, 10, 0, dynamics_plot_online)
-            actuator.setSleepTime(sleepTime)
-            actuator.run()
-            
-            monitor.running = 0
-            dynamics =  monitor.getDynamics()
-            actuator.close()
-            monitor.join()
-
-            data['x'].append(sleepTime)
-            data['y'].append(sum(dynamics['cpu']) / float(len(dynamics['cpu'])))
-
-            monitor.running = 0
-            monitor.join()
-
-            if not dynamics_plot_online:
-                plt.figure()
-                plt.plot(dynamics['time'], dynamics['cpu'])
-                plt.ylabel('CPU Load (%)')
-                plt.xlabel('Time [ms]')
-                plt.title('Dynamics_sleep_time_%fs.png' % (sleepTime))
-                plt.grid(True)
-                plt.savefig('Dynamics_sleep_time_%fs.png' % (sleepTime),dpi=100)
-                plt.close()
-        
[…]

### 2016-12-16 downloads/CPULoadGenerator/tests/Identification2.py

@@ -1,66 +1,0 @@
-#!/usr/bin/python
-
-#Authors: Gaetano Carlucci
-#         Giuseppe Cofano
-
-
-import time
-import json
-import matplotlib.pyplot as plt
-
-import sys
-import os
-sys.path.insert(0, os.path.dirname(os.path.abspath(__file__)) + '/../utils')
-
-from Monitor import MonitorThread
-from openLoopActuator import openLoopActuator
- 
-if __name__ == "__main__":
-   
-    ######################################################
-    #             IDENTIFICATION TEST                    #
-    ######################################################
-    # testing activities
-    # this test aims at characterizing the CPU
-    testing = 1
-    dynamics_plot_online = 0
-    if testing == 1:
-        sleepTimeSequence = [0.001, 0.005, 0.01, 0.02, 0.03, 0.08, 0.1, 0.15, 0.2, 0.3, 0.4, 0.5]
-        sleepTimeSequence = [0.001, 0.03, 0.02, 0.00, 0.04, 0.05, 0.03, 0.01, 0.2, 0.05, 0.01]
-        #sleepTimeSequence = [0.005,  0.02]
-
-        stepPeriod = 4
-        monitor = MonitorThread(0, 0.1)
-        monitor.start()
-        actuator = openLoopActuator(monitor, len(sleepTimeSequence) * stepPeriod, 0, dynamics_plot_online)
-        actuator.run_sequence(sleepTimeSequence)
-        
-        monitor.running = 0
-        dynamics =  monitor.getDynamics()
-        actuator.close()
-        monitor.join()
-        
-        with open('identification_data', 'w') as outfile:
-            json.dump(dynamics, outfile)
-    else:
-        with open('identification_data', 'r') as outfile:
-            dynamics = json.load(outfile)
-
-    plt.figure()
-    fig, ax1 = plt.subplots()
-    ax1.set_ylabel('CPU Load (%)', color='b')
-    ax1.plot(dynamics['time'], dynamics['cpu'], 'b-')
-    ax1.set_xlabel('Time [ms]')
-    ax1.grid(True)
-    for tl in ax1.get_yticklabels():
-        tl.set_color('b')
-    
-
-    ax2 = plt.twinx()
-    ax2.set_ylabel('Sleep Time Target(s)', color='r')
-    ax2.plot(dynamics['time'], dynamics['sleepTimeTarget'], 'r-')
-    ax2.grid(True)
[…]

### 2016-12-16 downloads/CPULoadGenerator/tests/PID.py

@@ -1,84 +1,0 @@
-#!/usr/bin/python
-
-#Authors: Gaetano Carlucci
-#         Giuseppe Cofano
-
-
-import json
-import matplotlib.pyplot as plt
-
-import sys
-import os
-sys.path.insert(0, os.path.dirname(os.path.abspath(__file__)) + '/../utils')
-
-from Monitor import MonitorThread
-from Controller import ControllerThread
-from closedLoopActuator import closedLoopActuator
-
-if __name__ == "__main__":
-   
-    ######################################################
-    #             PID TEST                               # 
-    ######################################################
-    # testing activities
-    # this test aims at characterizing the CPU
-    testing = 1
-    dynamics_plot_online = 0
-    if testing == 1:
-        cpuSequence = [0.8, 0.1, 0.30, 0.70, 0.40, 0.10, 0.20, 0.60, 0.20, 0.70]
-        #cpuSequence = [0.8, 0.1]
-        stepPeriod = 4
-
-        control = ControllerThread(0.1)
-        monitor = MonitorThread(0, 0.1)
-        actuator = closedLoopActuator(control, monitor, len(cpuSequence) * stepPeriod, 0, 1, dynamics_plot_online)
-
-        monitor.start()
-        control.start()
-        actuator.run_sequence(cpuSequence)
-        
-        actuator.close()
-        monitor.running = 0
-        control.running = 0
-        dynamics =  monitor.getDynamics()
-       
-        monitor.join()
-        control.join()
-        
-        stepPeriod = 4
-                
-        with open('pid_data', 'w') as outfile:
-            json.dump(dynamics, outfile)
-    else:
-        with open('pid_data', 'r') as outfile:
-            dynamics = json.load(outfile)
-
-    plt.figure()
-    fig, ax1 = plt.subplots()
-    ax1.set_ylabel('CPU Load (%)', color='b')
-    ax1.plot(dynamics['time'], dynamics['cpu'], 'b-')
-    ax1.set_xlabel('Time [ms]')
-    ax1.grid(True)
-    ax1.set_ylim([0, 100])
-    for tl in ax1.get_yticklabels():
-        tl.set_color('b')
-    
-    ax2 = plt.twinx()
-    ax2.set_ylabel('Input Load', color='r')
-    cpuInput = [ 100*(x) for x in dynamics['cpuTarget']]
[…]

### 2016-12-16 downloads/CPULoadGenerator/tests/feedForward.py

@@ -1,80 +1,0 @@
-#!/usr/bin/python
-
-#Authors: Gaetano Carlucci
-#         Giuseppe Cofano
-
-
-import json
-import matplotlib.pyplot as plt
-
-import sys
-import os
-sys.path.insert(0, os.path.dirname(os.path.abspath(__file__)) + '/../utils')
-
-from Monitor import MonitorThread
-from openLoopActuator import openLoopActuator
-
-period = 0.05 # actuation period  in seconds
-
-def cpu_model(cpu_target):
-    cpu_time = cpu_target * period
-    sleepTime = period - cpu_time
-    return sleepTime
-
-def inverse_cpu_model(sleepTime):
-    cpu_model = period - sleepTime
-    cpu_load =  cpu_model/period
-    return cpu_load
-
-if __name__ == "__main__":
-   
-    ######################################################
-    #             FEEDFORWARD TEST                    #
-    ######################################################
-    # testing activities
-    # this test aims at characterizing the CPU
-    testing = 1
-    dynamics_plot_online = 0
-    if testing == 1:
-        sleepTimeSequence = [0.001, 0.005, 0.01, 0.02, 0.03, 0.08, 0.1, 0.15, 0.2, 0.3, 0.4, 0.5]
-        cpuSequence = [0.1, 0.8, 0.30, 0.70, 0.40, 0.10, 0.20, 0.60, 0.20, 0.70]
-        #cpuSequence = [ 0.80]
-        #sleepTimeSequence = [0.001,  0.02]
-        sleepTimeSequence = [ cpu_model(x) for x in cpuSequence]
-        stepPeriod = 4
-        monitor = MonitorThread(0, 0.1)
-        actuator = openLoopActuator(monitor, len(sleepTimeSequence) * stepPeriod, 0, dynamics_plot_online)
-        monitor.start()
-        actuator.run_sequence(sleepTimeSequence)
-        
-        monitor.running = 0
-        dynamics =  monitor.getDynamics()
-        actuator.close()
-        monitor.join()
-        
-        with open('feed_forward_data', 'w') as outfile:
-            json.dump(dynamics, outfile)
-    else:
-        with open('feed_forward_data', 'r') as outfile:
-            dynamics = json.load(outfile)
-
-    plt.figure()
-    fig, ax1 = plt.subplots()
[…]

### 2016-12-16 downloads/CPULoadGenerator/utils/Controller.py

@@ -1,66 +1,0 @@
-#Authors: Gaetano Carlucci
-#         Giuseppe Cofano
-
-import threading
-import time
-
-class ControllerThread(threading.Thread):
-    """
-        Controls the CPU status
-    """
-    def __init__(self, interval, ki = None, kp = None):
-        self.running = 1;  # thread status
-        self.sampling_interval = interval
-        self.period = 0.1 # actuation period  in seconds
-        self.sleepTime = 0.02; # this is controller output: determines the sleep time to achieve the requested CPU load
-        self.alpha = 0.2; # filter coefficient
-        self.CT = 0.20;  # target CPU load should be provided as input 
-        self.cpu = 0;   # current CPU load returned from the Monitor thread
-        self.cpuPeriod = 0.03;
-        if ki is None:
-          self.ki = 0.2 #-0.005;   # integral constant of th PI regulator (the plant is an inverter)
-        if kp is None:
-          self.kp = 0.02  #1 #-0.05; # proportional constant of th PI regulator (the plant is an inverter)
-        self.int_err = 0;  # integral error
-        self.last_ts = time.time();  # last sampled time
-        super(ControllerThread, self).__init__()
-        
-    def getSleepTime(self):
-        return self.sleepTime
-
-    def cpu_model(self, cpu_period):
-      sleepTime = self.period - cpu_period
-      return sleepTime
-
-    def getCpuTarget(self):
-        return self.CT
-
-    def setCpu(self, cpu): 
-       self.cpu = self.alpha * cpu + (1 - self.alpha)*self.cpu # first order filter on the measurement samples
-
-    def getCpu(self): 
-       return self.cpu
-
-    def setCpuTarget(self, CT): 
-       self.CT = CT
-     
-    def run(self):
-        while self.running:
-          # ControllerThread has to have the same sampling interval as MonitorThread
-          time.sleep(self.sampling_interval)
-          self.err = self.CT - self.cpu*0.01  # computes the proportional error
-          ts = time.time()
-          
-          samp_int = ts - self.last_ts  # sample interval 
-          self.int_err = self.int_err + self.err*samp_int  # computes the integral error
-          self.last_ts = ts
-          self.cpuPeriod = self.kp*self.err  + self.ki*self.int_err
-
-          #anti wind up control
-          if self.cpuPeriod < 0:
-            self.cpuPeriod = 0
-            self.int_err = self.int_err - self.err*samp_int
[…]

### 2016-12-16 downloads/CPULoadGenerator/utils/Monitor.py

@@ -1,60 +1,0 @@
-#Authors: Gaetano Carlucci
-#         Giuseppe Cofano
-
-import os, psutil
-import threading
-import time
-
-class MonitorThread(threading.Thread):
-    """
-       Monitors the CPU status
-    """
-    def __init__(self, cpu_core, interval):
-        self.sampling_interval = interval; # sample time interval
-        self.sample = 0.5; # cpu load measurement sample
-        self.cpu = 0.5; # cpu load filtered
-        self.running = 1; # thread status
-        self.alpha = 1; # filter coefficient
-        self.sleepTimeTarget = 0.03
-        self.sleepTime = 0.03
-        self.cpuTarget = 0.5
-        self.cpu_core = cpu_core
-        self.dynamics = {"time":[], "cpu":[], "sleepTimeTarget":[], "cpuTarget":[],  "sleepTime":[],}
-        super(MonitorThread, self).__init__()
-        
-    def getCpuLoad(self):
-        return self.cpu
-
-    def setSleepTimeTarget(self, sleepTimeTarget):
-        self.sleepTimeTarget = sleepTimeTarget
-
-    def setSleepTime(self, sleepTime):
-        self.sleepTime = sleepTime
-
-    def setCPUTarget(self, cpuTarget):
-        self.cpuTarget = cpuTarget
-
-    def getDynamics(self):
-    	return self.dynamics
-        
-    def run(self):
-    	start_time = time.time()
-        p = psutil.Process(os.getpid())
-        try:
-            p.set_cpu_affinity([self.cpu_core]) #the process is forced to run only on the selected CPU
-        except AttributeError:
-            p.cpu_affinity([self.cpu_core])
-            
-        while self.running:
-            try:
-               self.sample = p.get_cpu_percent(self.sampling_interval)
-            except AttributeError:
-               self.sample = p.cpu_percent(self.sampling_interval)
-               
-            self.cpu = self.alpha * self.sample + (1 - self.alpha)*self.cpu # first order filter on the measurement samples
-            #self.cpu_log.append(self.cpu)
-            self.dynamics['time'].append(time.time() - start_time)
-            self.dynamics['cpu'].append(self.cpu)
-            self.dynamics['sleepTimeTarget'].append(self.sleepTimeTarget)
-            self.dynamics['sleepTime'].append(self.sleepTime)
-            self.dynamics['cpuTarget'].append(self.cpuTarget)

### 2016-12-16 downloads/CPULoadGenerator/utils/Plot.py

@@ -1,49 +1,0 @@
-#Authors: Gaetano Carlucci
-#         Giuseppe Cofano
-
-import time
-import matplotlib.pyplot as plt
-
-class realTimePlot():
-    """
-        Plots the CPU load
-    """
-    def __init__(self, duration, cpu, target):
-        plt.figure()
-        plt.axis([0, duration, 0, 100])
-        plt.ion()
-        plt.show()
-        plt.xlabel('Time(sec)')
-        plt.ylabel('%')
-        self.y_load = [0]
-        self.cpuT = target;
-        self.y_target = [0]
-        self.xdata = [0]       
-        self.line_load, = plt.plot(self.y_load)
-        self.line_target, = plt.plot(self.y_target)
-        if target != 0:
-            plt.legend([self.line_target, self.line_load], ["Target CPU", "CPU [%d] Load" % (cpu)], ncol=2)
-        else:
-            plt.legend([self.line_load], ["CPU [%d] Load" % (cpu)], ncol=1)
-        plt.grid(True)
-        self.ts_start = time.time()
-
-    def plotSample(self, sample, target):
-        p_x = time.time() - self.ts_start
-        p_load = sample
-        self.xdata.append(p_x)
-        if target != 0:
-            p_target = target 
-            self.y_target.append(p_target)
-            self.line_target.set_xdata(self.xdata)
-            self.line_target.set_ydata(self.y_target)
-        self.y_load.append(p_load)
-        self.line_load.set_xdata(self.xdata)
-        self.line_load.set_ydata(self.y_load)
-        plt.draw()
-
-    def close(self):
-        if self.cpuT != 0:
-            name = "%d%%-Target-Load" % (self.cpuT*100)+ ".jpg"
-            plt.savefig(name, dpi=100)
-        plt.close();

### 2016-12-16 downloads/CPULoadGenerator/utils/closedLoopActuator.py

@@ -1,72 +1,0 @@
-#Authors: Gaetano Carlucci
-#         Giuseppe Cofano
-
-import time
-
-from Plot import realTimePlot
-
-class closedLoopActuator():
-    """
-        Generates CPU load by tuning the sleep time
-    """
-    def __init__(self, controller, monitor, duration, cpu_core, target, plot):
-        self.controller = controller
-        self.monitor = monitor
-        self.duration = duration
-        self.plot = plot
-        self.target = target
-        self.controller.setCpu(self.monitor.getCpuLoad())
-        self.period = 0.05 # actuation period  in seconds
-        self.last_plot_time = time.time()
-        self.start_time = time.time()
-        if self.plot:
-            self.graph = realTimePlot(self.duration, cpu_core, target)
-
-    def generate_load(self, sleep_time):
-        interval = time.time() + self.cycle - sleep_time
-        # generates some getCpuLoad for interval seconds
-        while (time.time() < interval):
-            pr = 213123  # generates some load
-            pr * pr
-            pr = pr + 1
-        time.sleep(sleep_time) # controller actuation
-
-    def sendPlotSample(self):
-        if self.plot:
-            if (time.time() - self.last_plot_time) > 0.2:
-                self.graph.plotSample(self.controller.getCpu(), self.controller.getCpuTarget()*100)
-                self.last_plot_time = time.time()
-
-    def close(self):
-        if self.plot:
-            self.graph.close()
-
-    def generate_load(self, sleep_time):
-        interval = time.time() + self.period - sleep_time
-        # generates some getCpuLoad for interval seconds
-        while (time.time() < interval):
-            pr = 213123  # generates some load
-            pr * pr
-            pr = pr + 1
-
-        time.sleep(sleep_time)
-           
-    def run(self):
-        while (time.time() - self.start_time) <= self.duration:
-            self.controller.setCpu(self.monitor.getCpuLoad())
-            sleep_time = self.controller.getSleepTime()
-            self.generate_load(sleep_time)
-            self.sendPlotSample()
-        return sleep_time
-
-    def run_sequence(self, sequence):       
-        for cpuTarget in sequence:
-            stepPeriod = time.time() + 4
-            self.controller.setCpuTarget(cpuTarget)
-            self.monitor.setCPUTarget(cpuTarget)
-            while (time.time() < stepPeriod):
-                self.controller.setCpu(self.monitor.getCpuLoad())
[…]

### 2016-12-16 downloads/CPULoadGenerator/utils/openLoopActuator.py

@@ -1,62 +1,0 @@
-#Authors: Gaetano Carlucci
-#         Giuseppe Cofano
-
-import time
-
-from Plot import realTimePlot
-
-class openLoopActuator():
-    """
-        Generates CPU load by tuning the sleep time
-    """
-    def __init__(self, monitor, duration, cpu_core, plot):
-        self.sleep_time = 0.03
-        self.monitor = monitor
-        self.duration = duration
-        self.plot = plot
-        self.period = 0.05 # actuation period  in seconds
-        self.cpu_core = cpu_core
-        if self.plot:
-            self.graph = realTimePlot(self.duration, cpu_core, 0)
-
-    def setSleepTime(self, sleep_time):
-        self.sleep_time = sleep_time
-
-    def close(self):
-        if self.plot:
-            self.graph.close()
-
-    def checkSleepTime(self, sleep_time):
-        if sleep_time > self.period:
-            sleep_time = self.period
-        if sleep_time < 0:
-            sleep_time = 0
-        return sleep_time
-
-    def generate_load(self, sleep_time):
-        interval = time.time() + self.period - sleep_time
-        # generates some getCpuLoad for interval seconds
-        while (time.time() < interval):
-            pr = 213123  # generates some load
-            pr * pr
-            pr = pr + 1
-
-        time.sleep(sleep_time)
-
-    def sendPlotSample(self):
-        if self.plot:
-            self.graph.plotSample(self.monitor.getCpuLoad(), 0)
-
-    def run(self):
-        duration  = time.time() + self.duration
-        while (time.time() < duration):
-            self.generate_load(self.checkSleepTime(self.sleep_time))
-            self.sendPlotSample()
-
-    def run_sequence(self, sequence):       
-        for SleepTimeTarget in sequence:
-            stepPeriod = time.time() + 4
-            self.monitor.setSleepTimeTarget(SleepTimeTarget)
-            while (time.time() < stepPeriod):
-                self.generate_load(self.checkSleepTime(SleepTimeTarget))
-                self.sendPlotSample()

### 2016-12-16 ipfwd/ipfwd.c
Allow setting of accept all rule number

@@ -43,12 +43,12 @@ void control_ipfw(enum command cmd) {
   }
 }
 
-void set_probability(double probability) {
+void set_probability(int rule_number, double probability) {
   /* run the ipfw commands to set the new first rule probability
    */
   int status;
   pid_t pid;
-  char *rule_body = "-q add 1 prob 0.000 allow ip from any to any ";
+  char *rule_body = "-q add 00000 prob 0.000 allow ip from any to any ";
   char *set_rule_one;
 
   /* delete old rule */
@@ -69,7 +69,8 @@ void set_probability(double probability) {
       set_rule_one = malloc(sizeof(rule_body));
       snprintf(
           set_rule_one, strlen(rule_body),
-          "-q add 1 prob %0.3f allow ip from any to any", probability);
+          "-q add %d prob %0.3f allow ip from any to any", 
+          rule_number, probability);
 
       execl(ipfw_path, "ifpw", set_rule_one, NULL);
       exit(1); 
@@ -97,13 +98,18 @@ int main(int argc, char **argv) {
    * complete security. The probability is dynamic based on the current load.
    */
 
-  int status, i, num_cpus;
+  int status, i, num_cpus, rule_number;
   double slope, previous, prediction, error, probability;
   double loads[1], history[hist_size];
 
+  rule_number = 1;
   probability = 0.0;
   previous    = 100;
   num_cpus    = sysconf(_SC_NPROCESSORS_ONLN);
+
+  if (argc > 1)
+    rule_number = atoi(argv[1]);
+
   control_ipfw(ENABLE);
 
   while (1) {
@@ -124,7 +130,8 @@ int main(int argc, char **argv) {
       if (fabs(previous - history[0]) > sensitivity) {
         probability = sigmoid(history[0] / 100);
         previous    = history[0];
-        set_probability(probability);
+
+        set_probability(rule_number, probability);
       }
 
       printf(

### 2016-12-16 ipfwd/ipfwd.c
Allow setting of static acceptance probability

@@ -96,12 +96,21 @@ int main(int argc, char **argv) {
   /* monitors system load using uptime() to set probability of IPFW's first
    * rule allowing any packet. This rule increases performance by sacrificing
    * complete security. The probability is dynamic based on the current load.
+   *
+   * ipfwd [rule_number] [static_prob]
+   *   rule_number : int, rule number to use for accept all rule. rules before
+   *                 this will always be run. defaults to 1
+   *
+   *   static_prob : int, always use this probability for accept all rule,
+   *                 ignore current system load. 0 denotes no static_prob,
+   *                 defaults to 0
    */
 
-  int status, i, num_cpus, rule_number;
+  int status, i, num_cpus, rule_number, static_prob;
   double slope, previous, prediction, error, probability;
   double loads[1], history[hist_size];
 
+  static_prob = 0;
   rule_number = 1;
   probability = 0.0;
   previous    = 100;
@@ -109,6 +118,8 @@ int main(int argc, char **argv) {
 
   if (argc > 1)
     rule_number = atoi(argv[1]);
+  if (argc > 2)
+    static_prob = atoi(argv[2]);
 
   control_ipfw(ENABLE);
 
@@ -131,7 +142,10 @@ int main(int argc, char **argv) {
         probability = sigmoid(history[0] / 100);
         previous    = history[0];
 
-        set_probability(rule_number, probability);
+        if (static_prob)
+          set_probability(rule_number, static_prob);
+        else
+          set_probability(rule_number, probability);
       }
 
       printf(

### 2016-12-17 ipfwd/ipfwd.c
Finish adding static acceptance prob option

@@ -19,7 +19,8 @@ double sigmoid(double x) {
 void control_ipfw(enum command cmd) {
   /* enables or disables ipfw
    */
-  int enabled, on = 1, off = 0;
+  pid_t pid;
+  int status, enabled, on = 1, off = 0;
   size_t size = sizeof(enabled);
 
   sysctlbyname(sysctl_ipfw_v4, &enabled, &size, NULL, 0);
@@ -40,6 +41,18 @@ void control_ipfw(enum command cmd) {
       }
       /* do nothing if already off */
       break;
+    
+    case RESTART:
+      if ((pid = fork()) == 0) {
+        dup2(open("/dev/null", O_WRONLY), 1);
+        execl(service_path, "service", "ipfw", "restart", NULL);
+        exit(1); 
+      }
+      else {
+        waitpid(pid, &status, 0);
+        check(!status, "failed to restart ipfw during ipfwd startup"); 
+      }
+      break;
   }
 }
 
@@ -48,12 +61,22 @@ void set_probability(int rule_number, double probability) {
    */
   int status;
   pid_t pid;
-  char *rule_body = "-q add 00000 prob 0.000 allow ip from any to any ";
-  char *set_rule_one;
+  char *rule_body     = "-q add       prob 0.000 allow ip from any to any ";
+  char *rule_body_fmt = "-q add %d prob %0.3f allow ip from any to any";
+  char *dele_body     = "-q delete      ";
+  char *dele_body_fmt = "-q delete %d";
+  char *set_rule, *del_rule;
 
   /* delete old rule */
   if ((pid = fork()) == 0) {
-    execl(ipfw_path, "ipfw", "-q delete 1", NULL);
+    del_rule = malloc(sizeof(dele_body));
+    snprintf(
+        del_rule, strlen(dele_body), dele_body_fmt, 
+        rule_number);
+
+    dup2(open("/dev/null", O_WRONLY), 1);
+    dup2(open("/dev/null", O_WRONLY), 2);
+    execl(ipfw_path, "ipfw", del_rule, NULL);
     exit(1); 
   }
   else {
@@ -66,13 +89,13 @@ void set_probability(int rule_number, double probability) {
     if (probability > 0) {
       control_ipfw(ENABLE);
 
-      set_rule_one = malloc(sizeof(rule_body));
+      set_rule = malloc(sizeof(rule_body));
       snprintf(
-          set_rule_one, strlen(rule_body),
[…]

### 2016-12-17 ipfwd/ipfwd.h

@@ -12,6 +12,8 @@
 #include <string.h>
 #include <math.h>
 #include <time.h>
+#include <fcntl.h>
+#include <assert.h>
 #include <sys/types.h>
 #include <sys/wait.h>
 #include <sys/sysctl.h>
@@ -20,7 +22,7 @@
   if (! ret) { \
     fprintf(stderr, "error: %s\n", string); exit(1);}
 
-enum command {ENABLE, DISABLE};
+enum command {ENABLE, DISABLE, RESTART};
 
 const int sensitivity = 10;
 const int hist_size = 5;

### 2017-01-15 paper/PacketSamplingFirewalls.tex
Wrote outline for paper

@@
 \begin{itemize}
-    \item Problem
     \item Abstract
-    \item Subsampling
-    \item Randomized Subsampling
-    \item Dynamic Subsampling
-    \item Heuristic Based Rules
+    \item IPFWD
     \item IPFW
@@
 
-\textbf{Problem} \\
-Provide non-disruptive backend network monitoring to OneFS at 40Gb speeds.
+\textbf{Abstract} \\
+When using a firewall system like IPFW to detect threats, we can end up doing a
+lot of packet processing. This can negatively impact performance-sensitive
+systems such as storage nodes. This presentation describes a practical solution
+to this problem using a load-weighted probabilistic mechanism that allows a
+trade-off between perfect visibility of incoming packets and reduced impact to
+system load.
 
-\textbf{Abstract} \\
-The performance of FreeBSD’s IPFW can be increased to support 40Gb network
-interfaces without additional specialized hardware through dynamic
-probabilistic rule matching. The probability of rule matching will be based on
-current system load, allowing it to adapt to changing demands on the system.
-The efficacy of packet stream subsampling will also be explored.
+\textbf{IPFWD} \\
 
-\textbf{Subsampling} \\
-Ideally, all network traffic would be carefully inspected; this is impractical
-due to the limitations of CPU and bus speeds. When logging, it’s also limited
-by disk speed. Subsampling traffic provides a way take the temperature of the
-network and increase chances of detecting abnormal behavior; some monitoring is
-better than none. To provide full packet inspection to every packet on system,
-the network links would have to be split into 10Gb links, each passed through
-dedicated hardware.
+IPFWD is a daemon for FreeBSD that updates an early rule in
+IPFW that has a chance to accept any packet. The probability of acceptance is
+updated over time and dependent on the current system load. To account for
+this, IPFWD extends IPFW logs to include the likelihood an undetected policy
+violation occurred.
+
+This approach is based on the premise that firewall performance can be
+improved by reducing the number of rules applied to each packet. Supposing a
+whitelist policy, a firewall has to apply every rule to each packet before it's
+denied. With IPFWD, we have a chance to accept any packet early and skip any
+further computation. Test results shows this reduces the system resources
+required to handle the same amount of traffic.
+
+This is a shift in mindset from typical firewalls. Instead of enforcing every
+part of a security policy all the time, IPFWD enforces the policy some of the
+time and extrapolates from the violations it encounters. This provides insight
+into the violations that weren't detected. For this cost, you gain increased
+firewall performance and network throughput in resource bound systems. It's
+acceptable to allow a percentage of policy violations given that network
+traffic patterns are often repeated and the goal is detection and not immediate
+prevention. Preventative action may be taken later when resource requirements
+are lower.
+
+As an example, under heavy load IPFWD may immediately accept 40\% of packets.
+Supposing a port scan was initiated during this time and rules exist to block
+it, 60\% of the port scan would still be rejected and logged. Since the
+probability will fluctuate over time, IPFWD provides information in the IPFW
+logs to show the chance undetected violations occurred for each detected
+violation. IPFWD allows administrators to keep extensive rule sets that fully
+implement their security policy. Instead of having to simplify rule sets to
+increase performance, IPFWD balances policy enforcement and performance
+automatically. Under normal or light load, IPFWD will enforce the entire
+security policy 100\% of the time.
 

### 2017-01-22 paper/compile.sh
Clean up paper, working on ShareLatex now

@@ -1,3 +1,0 @@
-#!/bin/bash
-
-pdflatex -output-directory=/tmp/ PacketSamplingFirewalls.tex && cp /tmp/PacketSamplingFirewalls.pdf ~/Downloads

### 2017-01-23 paper/main.tex
Add LaTeX paper outline

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{IEEE Journal Template Example}
+\title{Probabilistic Policy Violation Detection with IPFWD}
 %
@@
 
-\author{Michael~Shell,~\IEEEmembership{Member,~IEEE,}
-        John~Doe,~\IEEEmembership{Fellow,~OSA,}
-        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
-\thanks{M. Shell is with the Department
-of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
-GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
-\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
-\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}
+\author{Dr. Phil Nelson and Austin Voecks% <-this % stops a space
+\thanks{placeholder}}
 
@@
 % The paper headers
-\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
+\markboth{March~2017}%
 {Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
@@
 
-% If you want to put a publisher's ID mark on the page you can do it like
-% this:
-%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
-% Remember, if you use this you must call \IEEEpubidadjcol in the second
-% column for its text to clear the IEEEpubid mark.
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
 % make the title area
@@
 %\boldmath
-\blindtext[1]
+When using a firewall system like IPFW to detect threats, we can end up doing a lot of packet processing.
+This can negatively impact performance-sensitive systems such as storage nodes.
+This paper describes a practical solution to this problem using a load-weighted probabilistic mechanism that allows a trade-off between perfect visibility of incoming packets and reduced impact to system load.
 \end{abstract}
-% IEEEtran.cls defaults to using nonbold math in the Abstract.
-% This preserves the distinction between vectors and scalars. However,
-% if the journal you are submitting to favors bold math in the abstract,
-% then you can use LaTeX's standard command \boldmath at the very start
-% of the abstract to achieve this. Many IEEE journals frown on math
-% in the abstract anyway.
 
@@
 \begin{IEEEkeywords}
-IEEEtran, journal, \LaTeX, paper, template.
+FreeBSD, IPFW, firewalls, performance.
 \end{IEEEkeywords}
@@
 \section{Introduction}
+
+IPFWD is a daemon for FreeBSD that updates an early rule in IPFW that has a chance to accept any packet.
+The probability of acceptance is updated over time and dependent on the current system load.
+To account for this, IPFWD extends IPFW logs to include the likelihood an undetected policy violation occurred.
+
+This approach is based on the premise that firewall performance can be improved by reducing the number of rules applied to each packet.
+Supposing a whitelist policy, a firewall has to apply every rule to each packet before it's denied.
+With IPFWD, we have a chance to accept any packet early and skip any further computation.
+Test results shows this reduces the system resources required to handle the same amount of traffic.
+
+This is a shift in mindset from typical firewalls.
+Instead of enforcing every part of a security policy all the time, IPFWD enforces the policy some of the time and extrapolates from the violations it encounters.
+This provides insight into the violations that weren't detected.
+For this cost, you gain increased firewall performance and network throughput in resource bound systems.
+It's acceptable to allow a percentage of policy violations given that network traffic patterns are often repeated and the goal is detection and not immediate prevention.
+Preventative action may be taken later when resource requirements are lower.
+
+As an example, under heavy load IPFWD may immediately accept 40\% of packets.
+Supposing a port scan was initiated during this time and rules exist to block it, 60\% of the port scan would still be rejected and logged.
+Since the probability will fluctuate over time, IPFWD provides information in the IPFW logs to show the chance undetected violations occurred for each detected violation.
+IPFWD allows administrators to keep extensive rule sets that fully implement their security policy.
+Instead of having to simplify rule sets to increase performance, IPFWD balances policy enforcement and performance automatically.
+Under normal or light load, IPFWD will enforce the entire security policy 100\% of the time.
+
+\section{IPFW}
+% Why was this firewall chosen?
+% First match wins rule system
+% Built in probabilistic match
 \blindtext
 
-\subsection{Subsection Heading Here}
+\section{IPFWD}
+% What does it do?
+% How does it work?
+% - Probabilistic packet matching
+% - Log extension
+% Explain rationale for it acting as it does
+% When should this system be used?
+% - High performance
+% - Resource conscious
 \blindtext
 
+\section{Security Implications}
+% Talk about security policy enforcement
+% What does it mean for security if your ignoring some of the packet?
+% How do you trust this system?
+% Talk about shift in mindset from "blocking attacks" to "detecting attacks"
+% - Why is this reasonable?
+% - Is this really different than traditional firewalls?
+\blindtext
+
+\section{Benchmarking Performance}
+% Explain different systems
+% Explain tools
+% Explain rationale for choosing tools
+\blindtext
+
+\section{Results}
+% Show results
+% Difference in results between CPU and throughput bound systems
+\blindtext
+
+\section{Related Work}
+% Cisco and Juniper routers
+% Machine learning firewalls
+\blindtext
+
+\section{Future Work}
+% Firewall rule reordering so we can "fail as early as possible"
+\blindtext
+
+\section{Conclusion}
+% What performance benefit do we get?
+% What security cost is there?
+% Why is this useful?
+% Why is this important?
+\blindtext
+
+% if have a single appendix:
+%\appendix[Proof of the Zonklar Equations]
+% or
+%\appendix  % for no appendix heading
+% do not use \section anymore after \appendix, only \section*
+% is possibly needed
+
+% use appendices with more than one appendix
+% then use \section to start each appendix
+% you must declare a \section before using any
+% \subsection or using \label (\appendices by itself
+% starts a section numbered zero.)
+%
+
+\appendices
+\section{Proof of the First Zonklar Equation}
+Some text for the appendix.
+
+% use section* for acknowledgement
+\section*{Acknowledgment}
+
+The authors would like to thank...
+
+% Can use something like this to put references on a page
+% by themselves when using endfloat and the captionsoff option.
+\ifCLASSOPTIONcaptionsoff
+  \newpage
+\fi
+
+% trigger a \newpage just before the given reference
+% number - used to balance the columns on the last page
+% adjust value as needed - may need to be readjusted if
+% the document is modified later
+%\IEEEtriggeratref{8}
+% The "triggered" command can be changed if desired:
+%\IEEEtriggercmd{\enlargethispage{-5in}}
+
+% references section
+
+% can use a bibliography generated by BibTeX as a .bbl file
+% BibTeX documentation can be easily obtained at:
+% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
+% The IEEEtran BibTeX style support page is at:
+% http://www.michaelshell.org/tex/ieeetran/bibtex/
+%\bibliographystyle{IEEEtran}
+% argument is your BibTeX string definitions and bibliography database(s)
+%\bibliography{IEEEabrv,../bib/paper}
+%
+% <OR> manually copy in the resultant .bbl file
+% set second argument of \begin to the number of references
+% (used to reserve space for the reference number labels box)
+\begin{thebibliography}{1}
+
+\bibitem{IEEEhowto:kopka}
+H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
+  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
+
+\end{thebibliography}
+
+% biography section
+% 
+% If you have an EPS/PDF photo (graphicx package needed) extra braces are
+% needed around the contents of the optional argument to biography to prevent
+% the LaTeX parser from getting confused when it sees the complicated
+% \includegraphics command within an optional argument. (You could create
+% your own custom macro containing the \includegraphics command to make things
+% simpler here.)
+%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
+% or if you just want to reserve a space for a photo:
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{picture}}]{John Doe}
+\blindtext
+\end{IEEEbiography}
+
+% You can push biographies down or up by placing
+% a \vfill before or after them. The appropriate
+% use of \vfill depends on what kind of text is
+% on the last page and whether or not the columns
+% are being equalized.
+
+%\vfill
+
+% Can be used to pull up biographies so that the bottom of the last one
+% is flush with the other column.
+%\enlargethispage{-5in}
+
+% that's all folks
+\end{document}
+
+% NOTES
+%==========================================================
 % needed in second column of first page if using \IEEEpubid
@@
 % stfloats package.
-
-\section{Conclusion}
-\blindtext
-
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
-
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
-
-\appendices
-\section{Proof of the First Zonklar Equation}
-Some text for the appendix.
-
-% use section* for acknowledgement
-\section*{Acknowledgment}
-
-The authors would like to thank...
-
-% Can use something like this to put references on a page
-% by themselves when using endfloat and the captionsoff option.
-\ifCLASSOPTIONcaptionsoff
-  \newpage
-\fi
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
-
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-\end{thebibliography}
-
-% biography section
-% 
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-%\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
-
-% that's all folks
-\end{document}

### 2017-02-01 paper/main.tex
Add table for test systems, explain tests

@@
 \usepackage{graphicx}
+\usepackage{array}
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 
@@
 
-IPFWD is a daemon for FreeBSD that updates an early rule in IPFW that has a chance to accept any packet.
-The probability of acceptance is updated over time and dependent on the current system load.
-To account for this, IPFWD extends IPFW logs to include the likelihood an undetected policy violation occurred.
+IPFWD is a daemon for FreeBSD that updates an early rule in IPFW that has a
+chance to accept any packet. The probability of acceptance is updated over time
+and dependent on the current system load. To account for this, IPFWD extends
+IPFW logs to include the likelihood an undetected policy violation occurred.
 
-This approach is based on the premise that firewall performance can be improved by reducing the number of rules applied to each packet.
-Supposing a whitelist policy, a firewall has to apply every rule to each packet before it's denied.
-With IPFWD, we have a chance to accept any packet early and skip any further computation.
-Test results shows this reduces the system resources required to handle the same amount of traffic.
+This approach is based on the premise that firewall performance can be improved
+by reducing the number of rules applied to each packet. Supposing a whitelist
+policy, a firewall has to apply every rule to each packet before it's denied.
+With IPFWD, we have a chance to accept any packet early and skip any further
+computation. Test results shows this reduces the system resources required to
+handle the same amount of traffic.
 
-This is a shift in mindset from typical firewalls.
-Instead of enforcing every part of a security policy all the time, IPFWD enforces the policy some of the time and extrapolates from the violations it encounters.
-This provides insight into the violations that weren't detected.
-For this cost, you gain increased firewall performance and network throughput in resource bound systems.
-It's acceptable to allow a percentage of policy violations given that network traffic patterns are often repeated and the goal is detection and not immediate prevention.
-Preventative action may be taken later when resource requirements are lower.
+This is a shift in mindset from typical firewalls. Instead of enforcing every
+part of a security policy all the time, IPFWD enforces the policy some of the
+time and extrapolates from the violations it encounters. This provides insight
+into the violations that weren't detected. For this cost, you gain increased
+firewall performance and network throughput in resource bound systems. It's
+acceptable to allow a percentage of policy violations given that network
+traffic patterns are often repeated and the goal is detection and not immediate
+prevention. Preventative action may be taken later when resource requirements
+are lower.
 
 As an example, under heavy load IPFWD may immediately accept 40\% of packets.
-Supposing a port scan was initiated during this time and rules exist to block it, 60\% of the port scan would still be rejected and logged.
-Since the probability will fluctuate over time, IPFWD provides information in the IPFW logs to show the chance undetected violations occurred for each detected violation.
-IPFWD allows administrators to keep extensive rule sets that fully implement their security policy.
-Instead of having to simplify rule sets to increase performance, IPFWD balances policy enforcement and performance automatically.
-Under normal or light load, IPFWD will enforce the entire security policy 100\% of the time.
+Supposing a port scan was initiated during this time and rules exist to block
+it, 60\% of the port scan would still be rejected and logged. Since the
+probability will fluctuate over time, IPFWD provides information in the IPFW
+logs to show the chance undetected violations occurred for each detected
+violation. IPFWD allows administrators to keep extensive rule sets that fully
+implement their security policy. Instead of having to simplify rule sets to
+increase performance, IPFWD balances policy enforcement and performance
+automatically. Under normal or light load, IPFWD will enforce the entire
+security policy 100\% of the time.
 
@@
 % Explain different systems
+Tests were run on three different systems and five different network cards.
+
+\begin{tabular}{ | l | L{19em} | } 
+ \hline
+ \multicolumn{2}{|c|}{Isilon OneFS 40Gb System} \\
+ \hline
+ \hline
+ OS  &  FreeBSD 11 Release \\
+ CPU &  8 x64 Intel\textsuperscript{\textregistered}\ Xeon\textsuperscript{\textregistered}\ CPU @ 2.20GHz \\
+ MEM &  64Gb \\
+ NIC &  Ethernet 40Gbase-T \\ 
+ \hline
+\end{tabular}\\\\
+
+\begin{tabular}{ | l | L{19em} | } 
+ \hline
+ \multicolumn{2}{|c|}{Isilon OneFS 10Gb System} \\
+ \hline
+ \hline
+ OS  &  FreeBSD 11 Release \\
+ CPU &  8 x64 Intel\textsuperscript{\textregistered}\ Xeon\textsuperscript{\textregistered}\ CPU @ 2.20GHz \\
+ MEM &  64Gb RAM \\
+ NIC &  Ethernet 10Gbase-T \\ 
+ \hline
+\end{tabular} \\\\
+
+\begin{tabular}{ | l | L{19em} | } 
+ \hline
+ \multicolumn{2}{|c|}{ WWU 10Gb System } \\
+ \hline
+ \hline
+ OS  &  FreeBSD 11 Release \\
+ CPU &  8 x64 Intel\textsuperscript{\textregistered}\ Core\texttrademark\ i7-2600 CPU @ 3.4GHz \\
+ MEM &  16Gb \\ 
+ NIC &  ??? \\ 
+ \hline
+\end{tabular} \\\\
+
+\begin{tabular}{ | l | L{19em} | } 
+ \hline
+ \multicolumn{2}{|c|}{ WWU 1Gb System } \\
+ \hline
+ \hline
+ OS  &  FreeBSD 11 Release \\
+ CPU &  8 x64 Intel\textsuperscript{\textregistered}\ Core\texttrademark\ i7-2600 CPU @ 3.4GHz \\
+ MEM &  16Gb \\
+ NIC &  Ethernet 1000baseT \\ 
+ \hline
+\end{tabular} \\\\
+
+\begin{tabular}{ | l | L{19em} | } 
+ \hline
+ \multicolumn{2}{|c|}{ Virtualized 10G System } \\
+ \hline
+ \hline
+ OS  &  FreeBSD 11 Release \\
+ CPU &  1 x64 Intel\textsuperscript{\textregistered}\ Xeon\textsuperscript{\textregistered}\ CPU @ 1.80GHz \\
+ MEM &  512Mb \\
+ NIC &  Ethernet 10Gbase-T \\ 
+ \hline
+\end{tabular} \\\\
+
+The systems tested all used 64bit FreeBSD 11 Release, with varying CPU type, CPU speed, RAM and network cards.
+Isilon and WWU systems were both physical hardware, while the Virtualized system was in a hypervisor environment.
+
+Virtualization makes it more difficult to have the same confidence in the results as with the physical machines.
+To account for this, tests on the Virtualized system were repeated additional times to account for the variability of the underlying physical hardware and contention between virtual machines.
+
+The WWU and Virtualized tests were conducted on stock FreeBSD 11 without additional network performance configuration applied.
+Isilon OneFS does contain additional network performance tuning and demonstrates the additional gains IPFWD can provide with careful FreeBSD 11 kernel configuration.
+
 % Explain tools
-% Explain rationale for choosing tools
-\blindtext
+Multiple network performance tools were explored but the final results were compiled using Netperf.
+Netperf was chosen because it was readily available on all test platforms, provides built-in CPU utilization measurement, and a wide range of test types.
+
+% Explain tests
+The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to preserve CPU cycles for IPFW.
+Raw network throughput and CPU performance is the target metric, which can be difficult to assess with TCP without performance tuning.
+(REFERENCE)
+
+All tests were conducted using the following steps:
+\begin{enumerate}
+    \item Machine 1
+    \begin{enumerate}
+        \item ipfw
+        \item netserver
+        \item CPU monitor
+    \end{enumerate}
+    \item Machine 2
+    \begin{enumerate}
+        \item ipfw
+        \item ipfwd
+        \item netperf
+        \item CPU monitor
+    \end{enumerate}
+\end{enumerate}
+
+All tests were run with minimal IPFW rule sets. Inituitively, larger
+production rule sets would benefit even further from IPFWD.
 

### 2017-02-05 paper/main.tex
More paper content, WWU 1G tests

@@
 
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
 % *** CITATION PACKAGES ***
@@
 % http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
 
@@
 
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
 % correct bad hyphenation here
@@
 \title{Probabilistic Policy Violation Detection with IPFWD}
-%
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-%
 
 \author{Dr. Phil Nelson and Austin Voecks% <-this % stops a space
-\thanks{placeholder}}
-
-% note the % following the last \IEEEmembership and also \thanks - 
-% these prevent an unwanted space from occurring between the last author name
-% and the end of the author line. i.e., if you had this:
-% 
-% \author{....lastname \thanks{...} \thanks{...} }
-%                     ^------------^------------^----Do not want these spaces!
-%
-% a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those "LaTeX things". For
-% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
-% "AB" then you have to do: "\textbf{A}\textbf{B}"
-% \thanks is no different in this regard, so shield the last } of each \thanks
-% that ends a line with a % and do not let a space in before the next \thanks.
-% Spaces after \IEEEmembership other than the last one are OK (and needed) as
-% you are supposed to have spaces between the names. For what it is worth,
-% this is a minor point as most people would not even notice if the said evil
-% space somehow managed to creep in.
+\thanks{Dell EMC Isilon}}
 
@@
 %\boldmath
-When using a firewall system like IPFW to detect threats, we can end up doing a lot of packet processing.
-This can negatively impact performance-sensitive systems such as storage nodes.
-This paper describes a practical solution to this problem using a load-weighted probabilistic mechanism that allows a trade-off between perfect visibility of incoming packets and reduced impact to system load.
+When using a firewall system like IPFW to detect threats, we can end up doing a
+lot of packet processing. This can negatively impact performance-sensitive
+systems such as storage nodes. This paper describes a practical solution to
+this problem using a load-weighted probabilistic mechanism that allows a
+trade-off between perfect visibility of incoming packets and reduced impact to
+system load.
+
 \end{abstract}
@@
 \section{IPFW}
+
+% Firewall basics
+Firewalls restrict the types of incoming and outgoing network traffic according
+to a security policy. Without a security policy, administrators cannot know
+what to allow and disallow. Typical firewalls enforce this policy by blocking
+packets that do not match the rules. Whitelist firewalls default action is to
+block all traffic, additional rules describe exceptions. Blacklist firewalls
+are the opposite in that by default they accept all traffic and additional
+rules describe packet types that are not allowed. Since predicting and
+blacklisting all possible attacks is infeasible, whitelist firewall rulesets
+are the accepted standard and the focus of this paper.
+
 % Why was this firewall chosen?
-% First match wins rule system
-% Built in probabilistic match
-\blindtext
+IPFW is a stateful firewall written for FreeBSD which supports both IPv4 and
+IPv6. It is comprised of several components: a kernel firewall filter rule
+processor and its integrated packet accounting facility, the logging facility,
+NAT, the dummynet traffic shaper, a forward facility, a bridge facility, and
+an ipstealth facility.
+
+IPFW was chosen for it's close integration with the FreeBSD operating system,
+in kernel packet filter, and general performance. Performance is dependent on
+system parameters and circumstances, but IPFW outperforms PF in stateful packet
+filtering. Stateful rules are more powerful and flexible than stateless rules,
+were no session information is maintained. Additionally, the capabilities of
+IPFW extend beyond packet filtering into source based routing, traffic shaping
+and more.
+
+% First match wins rule system and probabilistic matching
+Firewall rules are treated by IPFW in a first-match-wins fashion. IPFW also
+contains built-in functionality and rule syntax for probabilistic packet
+matching. From the documentation, this feature was intended for load balancing
+and other traffic shaping tasks. However it's the core of IPFWD's operation.
+
+This paper is not meant as a comparison between IPFW and PF; the functionality
+provided by IPFWD could be ported to PF and serves as a proof of concept for
+firewalls in general.
 
 \section{IPFWD}
+
 % What does it do?
+IPFWD increases firewall performance by reducing the average amount of work it
+takes to process a packet. Depending on the current system load, IPFWD
+increases or decreases the chance the IPFW will immediately accept a given
+packet. It does this by claiming and maintaining an early rule in the IPFW rule
+set of the following form: 'prob 0.000 allow ip from any to any'.
+
 % How does it work?
@@
 % - Log extension
+
 % Explain rationale for it acting as it does
+
 % When should this system be used?
@@
 \section{Security Implications}
+
 % Talk about security policy enforcement
+
 % What does it mean for security if your ignoring some of the packet?
+
 % How do you trust this system?
+
 % Talk about shift in mindset from "blocking attacks" to "detecting attacks"
@@
 \section{Benchmarking Performance}
+
 % Explain different systems
@@
 
+% Isilon 40G
 \begin{tabular}{ | l | L{19em} | } 
@@
 
+% Isilon 10G
 \begin{tabular}{ | l | L{19em} | } 
@@
 
+% WWU 10G
 \begin{tabular}{ | l | L{19em} | } 
@@
 
+% WWU 1G
 \begin{tabular}{ | l | L{19em} | } 
@@
 
+% Virtualized 10G
 \begin{tabular}{ | l | L{19em} | } 
@@
 
-The systems tested all used 64bit FreeBSD 11 Release, with varying CPU type, CPU speed, RAM and network cards.
-Isilon and WWU systems were both physical hardware, while the Virtualized system was in a hypervisor environment.
+The systems tested all used 64bit FreeBSD 11 Release, with varying CPU type,
+CPU speed, RAM and network cards. Isilon and WWU systems were both physical
+hardware, while the Virtualized system was in a hypervisor environment hosted
+by a third party service.
 
-Virtualization makes it more difficult to have the same confidence in the results as with the physical machines.
-To account for this, tests on the Virtualized system were repeated additional times to account for the variability of the underlying physical hardware and contention between virtual machines.
+Virtualization makes it more difficult to have the same confidence in the
+results as with the physical machines. Tests on the Virtualized system were
+repeated additional times to account for the variability of the underlying
+physical hardware and contention between virtual machines.
 
-The WWU and Virtualized tests were conducted on stock FreeBSD 11 without additional network performance configuration applied.
-Isilon OneFS does contain additional network performance tuning and demonstrates the additional gains IPFWD can provide with careful FreeBSD 11 kernel configuration.
+The WWU and Virtualized tests were conducted on stock FreeBSD 11 without
+additional network performance configuration applied. Isilon OneFS does contain
+additional network performance tuning and demonstrates the additional gains
+IPFWD can provide with careful FreeBSD 11 kernel configuration.
 
+\subsection{Netperf}
 % Explain tools
-Multiple network performance tools were explored but the final results were compiled using Netperf.
-Netperf was chosen because it was readily available on all test platforms, provides built-in CPU utilization measurement, and a wide range of test types.
+Multiple network performance tools were explored but the final results were
+compiled using Netperf. Netperf was chosen because it was readily available on
+all test platforms, provides built-in CPU utilization measurement, and a wide
+range of test types.
 
 % Explain tests
-The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to preserve CPU cycles for IPFW.
-Raw network throughput and CPU performance is the target metric, which can be difficult to assess with TCP without performance tuning.
-(REFERENCE)
+The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to
+preserve CPU cycles for IPFW. Raw network throughput and CPU performance is the
+target metric, which can be difficult to assess with TCP. (REFERENCE)
 
+Netperf is single threaded, which allows easier comparison of results between
+systems with different numbers of cores. Additionally, all tests were run while
+the system was idle to minimize contention over system resources.
+
+% Experimental method
 All tests were conducted using the following steps:
@@
 
-All tests were run with minimal IPFW rule sets. Inituitively, larger
-production rule sets would benefit even further from IPFWD.
+All tests were run with minimal IPFW rule sets in a range of 30 to 60 rules.
+Intuitively, larger production rule sets would benefit even further from IPFWD.
+
+CPU utilization measurement is the most difficult parameter to measure in these
+tests. There are many factors that can account for variability between tests
+and jitter in the results.
 
 \section{Results}
+
 % Show results
+
 % Difference in results between CPU and throughput bound systems
-\blindtext
+\subsection{Discussion}
+
+All the systems tested can be broadly categorized into two groups: CPU bound
+and NIC throughput bound. We can see that IPFWD affects performance positively
+but differently for both types.
+
+In CPU bound systems, IPFWD reduces the number of cycles it takes to process
+each packet, allowing more packets to be processed in the same time and
+increasing throughput. This is shown most clearly in the Virtualized system
+tests, which have the weakest CPU but 10Gb Ethernet NICs. This finding is also
+apparent in the 40Gb Isilon system.
+
+In NIC throughput bound systems, reducing the number of CPU cycles to process
+packets will not increase throughput. However, IPFWD still reduces CPU load
+which allows more cycles to be dedicated to other system processes.
 
@@
 % Cisco and Juniper routers
+
+Firewall decision trees? (Complete Redundancy Detection)
+
 % Machine learning firewalls
@@
 % Firewall rule reordering so we can "fail as early as possible"
-\blindtext
+Ideally, a whitelist firewall's rule set would be constructed so that the most
+commonly applied matches are reached earlier rather than later. IPFWD
+accomplishes this task by artificial early stopping. Firewall rule reordering
+is a could be another solution to this problem. Given logs of common network
+traffic for the system, statistical analysis could show which rules are matched
+more often than others. With this information, the rule set could be reordered
+so that the most commonly applied rules are earliest.
 
@@
 % What performance benefit do we get?
+
 % What security cost is there?
+
 % Why is this useful?
+
 % Why is this important?
+
 \blindtext

### 2017-02-06 paper/main.tex
Add graphs to paper

@@
 \usepackage{graphicx}
+\graphicspath{ {images/} }
 \usepackage{array}
@@
 % Explain different systems
-Tests were run on three different systems and five different network cards.
+Tests were run on three platforms and five different network cards, resulting
+in five different systems. The most important parameters to our tests are CPU,
+NIC, and ENV. Descriptions of each system are given in the tables below:\\
 
@@
  NIC &  Ethernet 40Gbase-T \\ 
+ ENV &  Physical \\ 
  \hline
@@
  NIC &  Ethernet 10Gbase-T \\ 
+ ENV &  Physical \\ 
  \hline
@@
  NIC &  ??? \\ 
+ ENV &  Physical \\ 
  \hline
@@
  NIC &  Ethernet 1000baseT \\ 
+ ENV &  Physical \\ 
  \hline
@@
  NIC &  Ethernet 10Gbase-T \\ 
+ ENV &  Hypervisor \\ 
  \hline
@@
   Multiple network performance tools were explored but the final results were
-compiled using Netperf. Netperf was chosen because it was readily available on
-all test platforms, provides built-in CPU utilization measurement, and a wide
-range of test types.
+  compiled using Netperf. Netperf was chosen because it was readily available
+  on all test platforms, provides built-in CPU utilization measurement, and a
+  wide range of test types.
 
@@
   The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to
-preserve CPU cycles for IPFW. Raw network throughput and CPU performance is the
-target metric, which can be difficult to assess with TCP. (REFERENCE)
+  preserve CPU cycles for IPFW. Raw network throughput and CPU performance is
+  the target metric, which can be difficult to assess with TCP.
 
   Netperf is single threaded, which allows easier comparison of results between
-systems with different numbers of cores. Additionally, all tests were run while
-the system was idle to minimize contention over system resources.
+  systems with different numbers of cores. Additionally, all tests were run
+  while the system was idle to minimize contention over system resources.
 
@@
   All tests were run with minimal IPFW rule sets in a range of 30 to 60 rules.
-Intuitively, larger production rule sets would benefit even further from IPFWD.
+  Intuitively, larger production rule sets would benefit even further from
+  IPFWD.
 
-CPU utilization measurement is the most difficult parameter to measure in these
-tests. There are many factors that can account for variability between tests
-and jitter in the results.
+  \subsection{CPU Measurement}
 
+  CPU utilization measurement is the most difficult parameter to measure in
+  these tests. There are many factors that can account for variability between
+  tests and jitter in the results. The Netperf manual describes these
+  challenges and the variety of techniques used on different systems.
+
+  Netperf was run in single threaded mode, which explains why throughput wasn't
+  closer to the maximum throughput for the NIC on each system. The FreeBSD
+  network stack allows multithreaded packet processing by utilizing multiple
+  queues. Since we're indirectly testing the performance of IPFW through
+  Netperf, there are two different CPU measurements to account for: kernel time
+  and user time. Since IPFW runs in the kernel, it's CPU time is classified as
+  kernel time.
+
+  Netperf was able to measure CPU utilization on the Virtualized and WWU
+  systems, but didn't work on the Isilon systems. Instead, ps was used on those
+  systems.  These tests are comparable since both measured total system CPU
+  utilization
+
+% Show results graphs, explain axises, etc
 \section{Results}
 
-% Show results
+  % throughput/CPU
+  \subsection{Throughput / CPU}
+  These graphs show network throughput divided by CPU utilization for each system. Higher values mean that the CPU time required to process packets is lower.
+
+  \begin{center}
+  \includegraphics[width=0.45\textwidth]{toc_isilon40}
+  \includegraphics[width=0.45\textwidth]{toc_isilon10}
+  \includegraphics[width=0.45\textwidth]{toc_wwu1}
+  \includegraphics[width=0.45\textwidth]{toc_virtual10}
+  \end{center}
+
+  %  throughput & CPU
+  \subsection{Throughput \& CPU}
+  These graphs show the interaction between CPU utilization and network throughput for each system with respect to IPFWD's acceptance probability.
+The dashed lines are CPU utilization, solid is network throughput in Mb/second.
+Higher acceptance probability means a greater chance that a given packet is immediately accepted without further processing.
+
+  \begin{center}
+  \includegraphics[width=0.45\textwidth]{isilon40_cpu}
+  \includegraphics[width=0.45\textwidth]{isilon10_cpu}
+  \includegraphics[width=0.45\textwidth]{wwu1_cpu}
+  \includegraphics[width=0.45\textwidth]{virtual10_cpu}
+  \end{center}
 
@@
   All the systems tested can be broadly categorized into two groups: CPU bound
-and NIC throughput bound. We can see that IPFWD affects performance positively
-but differently for both types.
+  and NIC throughput bound. We can see that IPFWD affects performance
+  positively but differently for both types.
 
@@
   increasing throughput. This is shown most clearly in the Virtualized system
-tests, which have the weakest CPU but 10Gb Ethernet NICs. This finding is also
-apparent in the 40Gb Isilon system.
+  tests, which have the weakest CPU but 10Gb Ethernet NICs. Regardless of test
+  type, the receiving system is always at 100\% CPU utilization. However, as
+  the probability to accept increases in IPFWD the throughput also increases.
+  This clearly shows that IPFWD decreases the CPU load required to process
+  incoming traffic.
 
@@
   packets will not increase throughput. However, IPFWD still reduces CPU load
-which allows more cycles to be dedicated to other system processes.
+  which allows more cycles to be dedicated to other system processes. This is
+  more difficult to see in the test results, particularly since the CPU
+  measurement tools are inherently imprecise and difficult to compare between
+  systems.
 
@@
 % Machine learning firewalls
-\blindtext
 
@@
 
-\blindtext
-
 % if have a single appendix:
@@
 \appendices
-\section{Proof of the First Zonklar Equation}
+\section{Additional Figures}
 Some text for the appendix.
@@
 
+% http://www.netperf.org/svn/netperf2/tags/netperf-2.7.0/doc/netperf.html#CPU-Utilization
+
 \end{thebibliography}
@@
 % or if you just want to reserve a space for a photo:
-
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{picture}}]{John Doe}
-\blindtext
-\end{IEEEbiography}
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-%\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
 

### 2017-02-06 paper/main.tex
Wrote IPFWD section, fix graphs

@@
 \author{Dr. Phil Nelson and Austin Voecks% <-this % stops a space
-\thanks{Dell EMC Isilon}}
+\thanks{In collaboration with Dell EMC Isilon}}
 
@@
 increases or decreases the chance the IPFW will immediately accept a given
-packet. It does this by claiming and maintaining an early rule in the IPFW rule
-set of the following form: 'prob 0.000 allow ip from any to any'.
+packet. It does this by adding and maintaining an early rule in the IPFW rule
+set of the following form: \verb|prob 0.000 allow ip from any to any|.
 
@@
 % - Probabilistic packet matching
+Since IPFW uses a first-match-wins rule system, this early rule will be
+encountered before the majority of the other rules in the rule set.  The
+probability given in this rule determines the chance that we skip processing
+all further rules, supposing that there hasn't been a match already.
+
+This is especially valuable for a whitelist firewall rule set dealing with
+large amounts of rejected traffic; normally each rejected packet has all rules
+applied to it while IPFW looks for a match.  When no matches are encountered,
+the default rule is employed and the packet is rejected.
+
+In a typical data center environment the vast majority of network traffic is
+valid and going to be accepted eventually. However, these rule sets are often
+complex and take increasing time to match a given packet depending on the
+length. For a rule set of length $n$, we can expect to a packet to encounter
+$n/2$ failed matches before matching the correct rule and being accepted.
+
+IPFWD improves the average number of failed matches before successful match
+from
+
+\[
+\frac{n}{2}
+\]
+
+to
+
+\[
+\Big(k \cdot \mathsf{P}(\textit{EA})\Big) + 
+\Big(\frac{n}{2} \cdot \mathsf{P}(\neg{\textit{EA}})\Big) 
+\]
+
+where $k$ is number of rules before the early acceptance rule and
+$\mathsf{P}(\textit{EA})$ is the probability that the early acceptance rule is
+matched.
+
+Removing the constant $k$, IPFWD reduces the number of failed matches before a
+successful match by a factor of
+
+\[
+\frac
+    {\frac{n}{2}}
+    {\frac{n}{2} \cdot \mathsf{P}(\neg{\textit{EA}})} 
+= \frac{1}{\mathsf{P}(\neg{\textit{EA})}}.
+\]
+
 % - Log extension
+However, while this method increases the speed the system processes packets, it
+creates a chance that packets that would have normally been rejected will get
+through the firewall. We'll refer to these types of packets as invald.
+
+To account for this, IPFWD extends the IPFW system logs to include information
+about the early accpetance probability. From this information, administrators
+can extrapolate the probability that additional invalid packets made it through
+the firewall for each that is detected. This is discussed further in the
+Security Implications section.
+
+% When should this system be used?
+IPFWD is not intended to be used on all types of systems. It's application
+instead most directly benefits the following types of systems:
+
+\begin{enumerate}
+    \item \textbf{High Network Performance}\\
+      The gains provided by IPFWD are only apparent in systems were network
+      throughput is matched by or out performs CPU speed. Systems that can
+      already easily handle fully processing each packet will not benefit by
+      reducing processing time. Given modern CPU clock speeds, systems with
+      network cards slower than 1 Gb/sec will likely not benefit either.
+
+    \item \textbf{Resource Conscious}\\
+      In line with high performance systems, systems with strict resource
+      restrictions can benefit by the reduced network related CPU usage.
+
+    \item \textbf{Complex Firewall Requirements}\\
+      IPFWD works by reducing the average number of rules that fail to match
+      before a packet is finished being processed. Very simple rulesets will
+      gain little from reducing the number of rules applied to each packet.
+
+    \item \textbf{Detection Based Security Policy}\\ 
+      By using IPFWD, administrators sacrifice the assurance that every invalid
+      packet will be rejected.  However, they are still provided information as
+      to whether some invalid packets were rejected, and that information can
+      be used to infer the existance of additional invalid packets.
+
+\end{enumerate}
 
 % Explain rationale for it acting as it does
-
-% When should this system be used?
-% - High performance
-% - Resource conscious
-\blindtext
+IPFWD is based on the assumptions that the majority of network traffic is
+valid, malicious events are rare, and
 
@@
 % - Is this really different than traditional firewalls?
-\blindtext
 
@@
   \begin{center}
-  \includegraphics[width=0.45\textwidth]{isilon40_cpu}
-  \includegraphics[width=0.45\textwidth]{isilon10_cpu}
-  \includegraphics[width=0.45\textwidth]{wwu1_cpu}
-  \includegraphics[width=0.45\textwidth]{virtual10_cpu}
+  \includegraphics[width=0.45\textwidth]{cpu_isilon40}
+  \includegraphics[width=0.45\textwidth]{cpu_isilon10}
+  \includegraphics[width=0.45\textwidth]{cpu_wwu1}
+  \includegraphics[width=0.45\textwidth]{cpu_virtual10}
   \end{center}
@@
 % Why is this important?
-
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
-
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
 

### 2017-02-14 paper/main.tex
Wrote IPFWD section, fix graphs
Reorganize tests, update main.tex

@@
 
-\author{Dr. Phil Nelson and Austin Voecks% <-this % stops a space
-\thanks{In collaboration with Dell EMC Isilon}}
+\author{Philip Nelson and Austin Voecks% <-this % stops a space
+\thanks{Supported in part by Dell EMC Isilon through a grant and hardware}}
 
@@
 chance to accept any packet. The probability of acceptance is updated over time
-and dependent on the current system load. To account for this, IPFWD extends
-IPFW logs to include the likelihood an undetected policy violation occurred.
+and dependent on the current system load. In high performance systems this results in an increase in network throughput.
 
@@
 policy, a firewall has to apply every rule to each packet before it's denied.
-With IPFWD, we have a chance to accept any packet early and skip any further
+With IPFWD, the system has a chance to accept any packet early and skip any further
 computation. Test results shows this reduces the system resources required to
@@
 % Firewall basics
+\subsection{Firewall Basics}
 Firewalls restrict the types of incoming and outgoing network traffic according
 to a security policy. Without a security policy, administrators cannot know
-what to allow and disallow. Typical firewalls enforce this policy by blocking
-packets that do not match the rules. Whitelist firewalls default action is to
+what to allow and disallow. Typical firewalls enforce the policy by blocking
+packets that do not match the rules. The default action for whitelist firewalls is to
 block all traffic, additional rules describe exceptions. Blacklist firewalls
@@
 % Why was this firewall chosen?
+\subsection{Choice of IPFW}
 IPFW is a stateful firewall written for FreeBSD which supports both IPv4 and
@@
 matching. From the documentation, this feature was intended for load balancing
-and other traffic shaping tasks. However it's the core of IPFWD's operation.
+and other traffic shaping tasks. It's the core of IPFWD's operation.
 
@@
 takes to process a packet. Depending on the current system load, IPFWD
-increases or decreases the chance the IPFW will immediately accept a given
-packet. It does this by adding and maintaining an early rule in the IPFW rule
+increases or decreases the chance the IPFW will accept a given
+packet early. It does this by adding and maintaining an early rule in the IPFW rule
 set of the following form: \verb|prob 0.000 allow ip from any to any|.
@@
 length. For a rule set of length $n$, we can expect to a packet to encounter
-$n/2$ failed matches before matching the correct rule and being accepted.
+$\frac{n}{2}$ failed matches before matching the correct rule and being accepted.
 
-IPFWD improves the average number of failed matches before successful match
+IPFWD improves the average number of failed matches before a successful match
 from
@@
 % - Log extension
-However, while this method increases the speed the system processes packets, it
+While this method increases the speed the system processes packets, it
 creates a chance that packets that would have normally been rejected will get
-through the firewall. We'll refer to these types of packets as invald.
+through the firewall. We'll refer to these types of packets as invald for the purpose of discussion.
 
-To account for this, IPFWD extends the IPFW system logs to include information
-about the early accpetance probability. From this information, administrators
+In order to provide visibility into this process, IPFWD extends the IPFW system logs to include information
+about the early acceptance probability. From this information, administrators
 can extrapolate the probability that additional invalid packets made it through
 the firewall for each that is detected. This is discussed further in the
-Security Implications section.
+\textit{Security Implications} section.
 
 % When should this system be used?
+\subsection{Application}
 IPFWD is not intended to be used on all types of systems. It's application
@@
 
+\subsection{Rationale}
 % Explain rationale for it acting as it does
 IPFWD is based on the assumptions that the majority of network traffic is
-valid, malicious events are rare, and
+valid, malicious events are rare, and TODO
 
@@
 
-% What does it mean for security if your ignoring some of the packet?
+% What does it mean for security if you're ignoring some of the packet?
 
@@
 
-The systems tested all used 64bit FreeBSD 11 Release, with varying CPU type,
+The systems tested all used x64 FreeBSD 11 Release, with varying CPU type,
 CPU speed, RAM and network cards. Isilon and WWU systems were both physical
@@
   systems.  These tests are comparable since both measured total system CPU
-  utilization
+  utilization TODO
 
@@
 Firewall decision trees? (Complete Redundancy Detection)
+TODO
 
@@
 so that the most commonly applied rules are earliest.
+TODO
 
 \section{Conclusion}
+TODO
 % What performance benefit do we get?
@@
 \section{Additional Figures}
-Some text for the appendix.
+TODO
 
@@
 \section*{Acknowledgment}
-
-The authors would like to thank...
+TODO
 

### 2017-02-14 ipfwd/ipfwd.c

@@ -159,20 +159,24 @@ int main(int argc, char **argv) {
     slope      = history[0] - history[1];
     prediction = history[0] + slope;
 
-    for (i = hist_size; i > 0; i--)
+    for (i = hist_size; i > 0; i--) {
       history[i] = history[i - 1];
+    }
 
     /* only take action when this load is different than the previous one */
     if (slope) {
 
       /* update probability if load has changed enough */
       if (fabs(previous - history[0]) > sensitivity) {
-        if (static_prob != -1.0)
-          probability = static_prob;
-        else
-          probability = sigmoid(history[0] / 100);
 
-        previous    = history[0];
+        if (static_prob != -1.0) {
+          probability = static_prob;
+        }
+        else {
+          probability = sigmoid(history[0] / 100);
+        }
+
+        previous = history[0];
         set_probability(rule_number, probability);
       }
 

### 2017-02-16 paper/main.tex
Refactor, remove IEEE notes, add bibliography, conclusion, related work

@@
 \documentclass[journal]{IEEEtran}
+
 \usepackage{blindtext}
+\usepackage[style=ieee,backend=biber]{biblatex}
+\usepackage[final]{microtype}
+\usepackage{csquotes}
+\usepackage[english]{babel} 
+\usepackage{array}
+
+\usepackage{hyperref}
+\addbibresource{references.bib}
+
 \usepackage{graphicx}
 \graphicspath{ {images/} }
-\usepackage{array}
+
 \newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at: 
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-%\ifCLASSOPTIONcaptionsoff
-%  \usepackage[nomarkers]{endfloat}
-% \let\MYoriglatexcaption\caption
-% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
-%\fi
-% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
-% This package may be useful when used in conjunction with IEEEtran.cls'
-% captionsoff option. Some IEEE journals/societies require that submissions
-% have lists of figures/tables at the end of the paper and that
-% figures/tables without any captions are placed on a page by themselves at
-% the end of the document. If needed, the draftcls IEEEtran class option or
-% \CLASSINPUTbaselinestretch interface can be used to increase the line
-% spacing as well. Be sure and use the nomarkers option of endfloat to
-% prevent endfloat from "marking" where the figures would have been placed
-% in the text. The two hack lines of code above are a slight modification of
-% that suggested by in the endfloat docs (section 8.3.1) to ensure that
-% the full captions always appear in the list of figures/tables - even if
-% the user used the short optional argument of \caption[]{}.
-% IEEE papers do not typically make use of \caption[]'s optional argument,
-% so this should not be an issue. A similar trick can be used to disable
-% captions of packages such as subfig.sty that lack options to turn off
-% the subcaptions:
-% For subfig.sty:
-% \let\MYorigsubfloat\subfloat
-% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
-% For subfigure.sty:
-% \let\MYorigsubfigure\subfigure
-% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
-% However, the above trick will not work if both optional arguments of
-% the \subfloat/subfig command are used. Furthermore, there needs to be a
-% description of each subfigure *somewhere* and endfloat does not add
-% subfigure captions to its list of figures. Thus, the best approach is to
-% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
-% and instead reference/explain all the subfigures within the main caption.
-% The latest version of endfloat.sty and its documentation can obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
-%
-% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
-% later in the document, say, to conditionally put the References on a 
-% page by themselves.
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+\providecommand*{\opt}[1]{\texttt{#1}}
+\providecommand*{\pkg}[1]{\textsf{#1}}
 
@@
 \author{Philip Nelson and Austin Voecks% <-this % stops a space
-\thanks{Supported in part by Dell EMC Isilon through a grant and hardware}}
+\thanks{Supported in part by Dell EMC Isilon through grant and hardware}}
 
@@
 
+% TODO rewrite
 \section{Introduction}
@@
   IPFWD is a daemon for FreeBSD that updates an early rule in IPFW that has a
-chance to accept any packet. The probability of acceptance is updated over time
-and dependent on the current system load. In high performance systems this results in an increase in network throughput.
+  chance to accept any packet. The probability of acceptance is updated over
+  time and dependent on the current system load. In high performance systems
+  this results in an increase in network throughput.
 
-This approach is based on the premise that firewall performance can be improved
-by reducing the number of rules applied to each packet. Supposing a whitelist
-policy, a firewall has to apply every rule to each packet before it's denied.
-With IPFWD, the system has a chance to accept any packet early and skip any further
-computation. Test results shows this reduces the system resources required to
-handle the same amount of traffic.
+  This approach is based on the premise that firewall performance can be
+  improved by reducing the number of rules applied to each packet. Supposing a
+  whitelist policy, a firewall has to apply every rule to each packet before
+  it's denied.  With IPFWD, the system has a chance to accept any packet early
+  and skip any further computation. Test results shows this reduces the system
+  resources required to handle the same amount of traffic.
 
-This is a shift in mindset from typical firewalls. Instead of enforcing every
-part of a security policy all the time, IPFWD enforces the policy some of the
-time and extrapolates from the violations it encounters. This provides insight
-into the violations that weren't detected. For this cost, you gain increased
-firewall performance and network throughput in resource bound systems. It's
-acceptable to allow a percentage of policy violations given that network
-traffic patterns are often repeated and the goal is detection and not immediate
-prevention. Preventative action may be taken later when resource requirements
-are lower.
+  This is a shift in mindset from typical firewalls \cite{networksecurity}.
+  Instead of enforcing every part of a security policy all the time, IPFWD
+  enforces the policy some of the time and extrapolates from the violations it
+  encounters.  This provides insight into the violations that weren't detected.
+  For this cost, you gain increased firewall performance and network throughput
+  in resource bound systems. It's acceptable to allow a percentage of policy
+  violations given that network traffic patterns are often repeated and the
+  goal is detection and not immediate prevention. Preventative action may be
+  taken later when resource requirements are lower.
 
@@
   \subsection{Firewall Basics}
-Firewalls restrict the types of incoming and outgoing network traffic according
-to a security policy. Without a security policy, administrators cannot know
-what to allow and disallow. Typical firewalls enforce the policy by blocking
-packets that do not match the rules. The default action for whitelist firewalls is to
-block all traffic, additional rules describe exceptions. Blacklist firewalls
-are the opposite in that by default they accept all traffic and additional
-rules describe packet types that are not allowed. Since predicting and
-blacklisting all possible attacks is infeasible, whitelist firewall rulesets
-are the accepted standard and the focus of this paper.
+
+    Firewalls restrict the types of incoming and outgoing network traffic
+    according to a security policy. Without a security policy, administrators
+    cannot know what to allow and disallow. Typical firewalls enforce the
+    policy by blocking packets that do not match the rules. The default action
+    for whitelist firewalls is to block all traffic, additional rules describe
+    exceptions. Blacklist firewalls are the opposite in that by default they
+    accept all traffic and additional rules describe packet types that are not
+    allowed. Since predicting and blacklisting all possible attacks is
+    infeasible, whitelist firewall rulesets are the accepted standard and the
+    focus of this paper.
 
@@
   \subsection{Choice of IPFW}
-IPFW is a stateful firewall written for FreeBSD which supports both IPv4 and
-IPv6. It is comprised of several components: a kernel firewall filter rule
-processor and its integrated packet accounting facility, the logging facility,
-NAT, the dummynet traffic shaper, a forward facility, a bridge facility, and
-an ipstealth facility.
 
-IPFW was chosen for it's close integration with the FreeBSD operating system,
-in kernel packet filter, and general performance. Performance is dependent on
-system parameters and circumstances, but IPFW outperforms PF in stateful packet
-filtering. Stateful rules are more powerful and flexible than stateless rules,
-were no session information is maintained. Additionally, the capabilities of
-IPFW extend beyond packet filtering into source based routing, traffic shaping
+    IPFW is a stateful firewall written for FreeBSD which supports both IPv4
+    and IPv6. It is comprised of several components: a kernel firewall filter
+    rule processor and its integrated packet accounting facility, the logging
+    facility, NAT, the dummynet traffic shaper, a forward facility, a bridge
+    facility, and an ipstealth facility \cite{freebsdhandbook}.
+
+    IPFW was chosen for it's close integration with the FreeBSD operating
+    system, in kernel packet filter, and general performance. Performance is
+    dependent on system parameters and circumstances, but IPFW outperforms PF
+    in stateful packet filtering \cite{ipfwvspf}. Stateful rules are more
+    powerful and flexible than stateless rules, were no session information is
+    maintained \cite{networksecurity}.  Additionally, the capabilities of IPFW
+    extend beyond packet filtering into source based routing, traffic shaping
     and more.
@@
     contains built-in functionality and rule syntax for probabilistic packet
-matching. From the documentation, this feature was intended for load balancing
-and other traffic shaping tasks. It's the core of IPFWD's operation.
+    matching. From the documentation, this feature was intended for load
+    balancing and other traffic shaping tasks. It's the core of IPFWD's
+    operation.
 
-This paper is not meant as a comparison between IPFW and PF; the functionality
-provided by IPFWD could be ported to PF and serves as a proof of concept for
-firewalls in general.
+    This paper is not meant as a comparison between IPFW and PF; the
+    functionality provided by IPFWD could be ported to PF and serves as a proof
+    of concept for firewalls in general.
 
@@
   % What does it do?
-IPFWD increases firewall performance by reducing the average amount of work it
-takes to process a packet. Depending on the current system load, IPFWD
-increases or decreases the chance the IPFW will accept a given
-packet early. It does this by adding and maintaining an early rule in the IPFW rule
-set of the following form: \verb|prob 0.000 allow ip from any to any|.
+  IPFWD increases firewall performance by reducing the average amount of work
+  it takes to process a packet. Depending on the current system load, IPFWD
+  increases or decreases the chance the IPFW will accept a given packet early.
+  It does this by adding and maintaining an early rule in the IPFW rule set of
+  the following form: \verb|prob 0.000 allow ip from any to any|.
 
@@
   % - Probabilistic packet matching
+  IPFWD's behavior is based on the assumption that the majority of network
+  traffic is valid, malicious events are rare \cite{networktrafficanalysis},
+  and it's sufficient to be notified of security policy violations instead of
+  always stopping them \cite{TODO}.
+
   Since IPFW uses a first-match-wins rule system, this early rule will be
@@
   This is especially valuable for a whitelist firewall rule set dealing with
-large amounts of rejected traffic; normally each rejected packet has all rules
-applied to it while IPFW looks for a match.  When no matches are encountered,
-the default rule is employed and the packet is rejected.
+  large amounts of rejected traffic; normally each rejected packet has all
+  rules applied to it while IPFW looks for a match.  When no matches are
+  encountered, the default rule is employed and the packet is rejected.
 
   In a typical data center environment the vast majority of network traffic is
-valid and going to be accepted eventually. However, these rule sets are often
-complex and take increasing time to match a given packet depending on the
-length. For a rule set of length $n$, we can expect to a packet to encounter
-$\frac{n}{2}$ failed matches before matching the correct rule and being accepted.
+  valid and going to be accepted eventually \cite{TODO}. However, these rule
+  sets are often complex and take increasing time to match a given packet
+  depending on the length. For a rule set of length $n$, we can expect to a
+  packet to encounter $\frac{n}{2}$ failed matches before matching the correct
+  rule and being accepted.
 
@@
   where $k$ is number of rules before the early acceptance rule and
-$\mathsf{P}(\textit{EA})$ is the probability that the early acceptance rule is
-matched.
+  $\mathsf{P}(\textit{EA})$ is the probability that the early acceptance rule
+  is matched.
 
-Removing the constant $k$, IPFWD reduces the number of failed matches before a
-successful match by a factor of
+  Removing the constant $k$, IPFWD reduces the number of failed matches before
+  a successful match by a factor of
 
@@
   creates a chance that packets that would have normally been rejected will get
-through the firewall. We'll refer to these types of packets as invald for the purpose of discussion.
+  through the firewall. We'll refer to these types of packets as invalid for
+  the purpose of discussion.
 
-In order to provide visibility into this process, IPFWD extends the IPFW system logs to include information
-about the early acceptance probability. From this information, administrators
-can extrapolate the probability that additional invalid packets made it through
-the firewall for each that is detected. This is discussed further in the
-\textit{Security Implications} section.
+  In order to provide visibility into this process, IPFWD extends the IPFW
+  system logs to include information about the early acceptance probability.
+  From this information, administrators can extrapolate the probability that
+  additional invalid packets made it through the firewall for each that is
+  detected. This is discussed further in the \textit{Security Implications}
+  section.
 
@@
   \subsection{Application}
+
     IPFWD is not intended to be used on all types of systems. It's application
@@
     \begin{enumerate}
-    \item \textbf{High Network Performance}\\
+      \item \textbf{High Network Performance}
+
         The gains provided by IPFWD are only apparent in systems were network
@@
 
-    \item \textbf{Resource Conscious}\\
+      \item \textbf{Resource Conscious}
+
         In line with high performance systems, systems with strict resource
@@
 
-    \item \textbf{Complex Firewall Requirements}\\
+      \item \textbf{Complex Firewall Requirements}
+
         IPFWD works by reducing the average number of rules that fail to match
@@
 
-    \item \textbf{Detection Based Security Policy}\\ 
-      By using IPFWD, administrators sacrifice the assurance that every invalid
-      packet will be rejected.  However, they are still provided information as
-      to whether some invalid packets were rejected, and that information can
-      be used to infer the existance of additional invalid packets.
+      \item \textbf{Detection Based Security Policy}
+
+        By using IPFWD, administrators sacrifice the assurance that every
+        invalid packet will be rejected.  However, they are still provided
+        information as to whether some invalid packets were rejected, and that
+        information can be used to infer the existence of additional invalid
+        packets.
 
     \end{enumerate}
-
-\subsection{Rationale}
-% Explain rationale for it acting as it does
-IPFWD is based on the assumptions that the majority of network traffic is
-valid, malicious events are rare, and TODO
 
@@
   % Talk about security policy enforcement
+  Security policies define what is and is not allowed in an organization
+  \cite{TODO}. At the network and host level, this is implemented in part by a
+  firewall rule set. A policy violation on an internal network is more serious
+  than an external network. Internet facing hosts can expect to be scanned,
+  experience network anomalies and be attacked by third party agents
+  semi-regularly.
+
+  However when this happens on an internal network, there is a strong
+  possibility something else is going wrong. The usage of IPFWD is predicated
+  on the assumption that it's sufficient to be notified of these events so
+  further action can be taken later than to always immediately stop them.
 
   % What does it mean for security if you're ignoring some of the packet?
+  Allowing some percentage of network packets through the firewall without
+  inspection does mean that attacks that would have normally been blocked could
+  succeed. In addition to performance improvements, the following factors help
+  to account for this risk:
+
+  \begin{enumerate}
+
+    % Bad things are often repeated
+    \item \textbf{Repetition}
+
+      Malicious network activity and network errors are likely to be repeated
+      over time \cite{TODO}. The longer the activity persists, the more likely
+      it is to be detected by IPFW even when IPFWD is allowing a high
+      percentage of the network traffic through unchecked.
+
+    \item \textbf{Disruption}
+
+      Though some malicious network activity may only consist of a single
+      packet, others will require multiple packets. Since IPFW works on a
+      per-packet basis, malicious network activity will experience
+      approximately $m\%$ packet loss where $m$ is IPFWD's probability of early
+      acceptance. The exact effect this would have on malware or reconnaissance
+      would depend, but generally will increase the chance they fail or produce
+      invalid results.
+
+  \end{enumerate}
 
   % How do you trust this system?
+  Knowledge of a network's traffic patterns and the additional log information
+  provided by IPFWD can help. It allows administrators to extrapolate the
+  existence of undetected network events through simple statistical models.
 
@@
   % - Is this really different than traditional firewalls?
+  TODO
 
+  % example application
+  \subsection{Example Scenario}
+
+    Suppose each host in a data center's internal network is running IPFWD on
+    each host, average CPU usage is $75\%$, and have rule sets blocking common
+    network attacks. One of the hosts has been compromised and begins
+    conducting network reconnaissance by port scanning the other hosts on the
+    network.
+
+    Firewall rules are in place to block host to host communication $m\%$ of
+    the ports being scanned. This means that $1 - m\%$ of the scan would be
+    allowed regardless of IPFWD's action. This percentage represents valid
+    host-to-host communication as determined by the security policy.
+
+    IPFWD would allow $75\%$ of the policy violating port scan through to the
+    host, and $75\%$ of the replies back to the compromised host. This results
+    in $75\% * 75\% = 56.25\%$ of the port scan that should have been stopped
+    succeeding.  However, the other $1 - 56.25 = 43.75\%$ of the port scan was
+    blocked and logged, alerting system administrators to the presence of
+    malicious network activity so further action can be taken.
+
+  % describe difference between reply and no reply violations
+  \subsection{Security Cost}
+
+    The chance of missing a security policy violation depends on the type of
+    traffic. The following situations exist when both incoming and outgoing
+    firewall rules are employed:
+
+    \begin{enumerate}
+
+      % only one chance to catch
+      \item \textbf{Incoming or Outgoing + No Reply}
+
+        Here, the firewall only has one chance to catch the policy violation.
+        So if the early acceptance probability is $m$, then $m\%$ of the policy
+        violating packets will be allowed.
+
+      % two chances to catch 
+      \item \textbf{Incoming or Outgoing + Reply}
+
+        Here, the firewall has two chances to catch the policy violation.  So
+        if the early acceptance probability is $m$, then $m'\%$ of the policy
+        violating packets will be allowed where $m' = m * m$.
+
+    \end{enumerate}
+
+% Explain different systems and tests
 \section{Benchmarking Performance}
 
-% Explain different systems
   Tests were run on three platforms and five different network cards, resulting
-in five different systems. The most important parameters to our tests are CPU,
-NIC, and ENV. Descriptions of each system are given in the tables below:\\
+  in five different systems. The most important parameters to our tests are
+  CPU, NIC, and ENV. Descriptions of each system are given in the tables
+  below:\\
 
@@
 
-% WWU 10G
-\begin{tabular}{ | l | L{19em} | } 
- \hline
- \multicolumn{2}{|c|}{ WWU 10Gb System } \\
- \hline
- \hline
- OS  &  FreeBSD 11 Release \\
- CPU &  8 x64 Intel\textsuperscript{\textregistered}\ Core\texttrademark\ i7-2600 CPU @ 3.4GHz \\
- MEM &  16Gb \\ 
- NIC &  ??? \\ 
- ENV &  Physical \\ 
- \hline
-\end{tabular} \\\\
-
   % WWU 1G
@@
 
-% Virtualized 10G
+  % Virtualised 10G
   \begin{tabular}{ | l | L{19em} | } 
     \hline
- \multicolumn{2}{|c|}{ Virtualized 10G System } \\
+    \multicolumn{2}{|c|}{ Virtualised 10G System } \\
     \hline
@@
 
-The systems tested all used x64 FreeBSD 11 Release, with varying CPU type,
-CPU speed, RAM and network cards. Isilon and WWU systems were both physical
-hardware, while the Virtualized system was in a hypervisor environment hosted
+  The systems were tested on x64 FreeBSD 11 Release, with varying CPU type, CPU
+  speed, RAM and network cards. Isilon and WWU systems were both physical
+  hardware, while the Virtualised system was in a hypervisor environment hosted
   by a third party service.
 
-Virtualization makes it more difficult to have the same confidence in the
-results as with the physical machines. Tests on the Virtualized system were
+  Virtualisation makes it more difficult to have the same confidence in the
+  results as with the physical machines. Tests on the Virtualised system were
   repeated additional times to account for the variability of the underlying
@@
 
-The WWU and Virtualized tests were conducted on stock FreeBSD 11 without
-additional network performance configuration applied. Isilon OneFS does contain
-additional network performance tuning and demonstrates the additional gains
-IPFWD can provide with careful FreeBSD 11 kernel configuration.
+  The WWU and Virtualised tests were conducted on stock FreeBSD 11 without
+  additional network performance configuration applied. Isilon OneFS does
+  contain additional network performance tuning and demonstrates the additional
+  gains IPFWD can provide with careful FreeBSD 11 kernel configuration.
 
   \subsection{Netperf}
+
     % Explain tools
@@
     on all test platforms, provides built-in CPU utilization measurement, and a
-  wide range of test types.
+    wide range of test types \cite{netperf}.
 
@@
     preserve CPU cycles for IPFW. Raw network throughput and CPU performance is
-  the target metric, which can be difficult to assess with TCP.
+    the target metric, which can be difficult to assess with TCP
+    \cite{TODO}.
 
-  Netperf is single threaded, which allows easier comparison of results between
-  systems with different numbers of cores. Additionally, all tests were run
-  while the system was idle to minimize contention over system resources.
+    Netperf is single threaded, which allows easier comparison of results
+    between systems with different numbers of cores. Additionally, all tests
+    were run while the system was idle to minimize contention over system
+    resources.
 
@@
         \item Machine 1
-      \begin{enumerate}
-          \item ipfw
-          \item netserver
-          \item CPU monitor
-      \end{enumerate}
+        \begin{itemize}
+          \item \texttt{ipfw}
+          \item \texttt{netserver}
+          \item \texttt{CPU monitor}
+        \end{itemize}
         \item Machine 2
-      \begin{enumerate}
-          \item ipfw
-          \item ipfwd
-          \item netperf
-          \item CPU monitor
-      \end{enumerate}
+        \begin{itemize}
+          \item \texttt{ipfw}
+          \item \texttt{ipfwd}
+          \item \texttt{netperf}
+          \item \texttt{CPU monitor}
+        \end{itemize}
     \end{enumerate}
 
-  All tests were run with minimal IPFW rule sets in a range of 30 to 60 rules.
-  Intuitively, larger production rule sets would benefit even further from
-  IPFWD.
+    All tests were run with minimal IPFW rule sets in a range of 30 to 60 rules
+    depending on the environment and network card configuration. Larger
+    production rule sets would benefit even further from IPFWD.
 
@@
     CPU utilization measurement is the most difficult parameter to measure in
-  these tests. There are many factors that can account for variability between
-  tests and jitter in the results. The Netperf manual describes these
+    these tests. There are many factors that can account for variability
+    between tests and jitter in the results. The Netperf manual describes these
     challenges and the variety of techniques used on different systems.
+    \cite{netperf}
 
-  Netperf was run in single threaded mode, which explains why throughput wasn't
-  closer to the maximum throughput for the NIC on each system. The FreeBSD
-  network stack allows multithreaded packet processing by utilizing multiple
-  queues. Since we're indirectly testing the performance of IPFW through
-  Netperf, there are two different CPU measurements to account for: kernel time
-  and user time. Since IPFW runs in the kernel, it's CPU time is classified as
-  kernel time.
+    Netperf was run in single threaded mode, which explains why throughput
+    wasn't closer to the maximum throughput for the NIC on each system. The
+    FreeBSD network stack allows multithreaded packet processing by utilizing
+    multiple queues \cite{freebsdhandbook}. Since we're indirectly testing the
+    performance of IPFW through Netperf, there are two different CPU
+    measurements to account for: kernel time and user time. Since IPFW runs in
+    the kernel, it's CPU time is classified as kernel time.
 
     Netperf was able to measure CPU utilization on the Virtualized and WWU
-  systems, but didn't work on the Isilon systems. Instead, ps was used on those
-  systems.  These tests are comparable since both measured total system CPU
-  utilization TODO
+    systems, but didn't work on the Isilon systems. Instead, \textit{ps} was
+    used. These tests are comparable since both measured user CPU utilization,
+    not kernel CPU utilization.
+
+    In systems where the throughput of the NIC exceeds what the CPU can supply,
+    CPU utilization is high and throughput is lower than the NIC maximum (eg:
+    Virtualised system). Conversely, when CPU availablility exceeds NIC
+    throughput, throughput is high and CPU utilization is low (eg: WWU 1G
+    system).
 
@@
 
+  %  throughput & CPU
+  \subsection{Throughput \& CPU}
+
+    These graphs show the interaction between CPU utilization and network
+    throughput for each system with respect to IPFWD's early acceptance
+    probability. The dashed lines are CPU utilization, solid is network
+    throughput in Mb/second. Higher acceptance probability means a greater
+    chance that a given packet is immediately accepted without further
+    processing.
+
+    The upward trend in throughput on the x-axis shows that as the early
+    acceptance probability increases, so does throughput. The CPU usage shows
+    that this throughput is gained without additional CPU utilization, meaning
+    the system is spending fewer cycles to process each packet on average.
+
+    The lack of any trend in the WWU 1G graph is expected and addressed in the
+    \textit{Discussion} section.
+
+    \begin{center}
+      \includegraphics[width=0.45\textwidth]{cpu_isilon40}
+      \includegraphics[width=0.45\textwidth]{cpu_isilon10}
+      \includegraphics[width=0.45\textwidth]{cpu_wwu1}
+      \includegraphics[width=0.45\textwidth]{cpu_virtual10}
+    \end{center}
+
   % throughput/CPU
   \subsection{Throughput / CPU}
-  These graphs show network throughput divided by CPU utilization for each system. Higher values mean that the CPU time required to process packets is lower.
+
+    These graphs show network the ratio of throughput to CPU utilization for
+    each system where $ratio(x) = throughput(x) / cpu(x)$. The data being
+    presented here is the same as the previous graphs, shown differently. The
+    upward trend over the x-axis shows that as the probability of early
+    acceptance increases, the system is spending less CPU time to process
+    packets on average.
+
+    The dashed line shows the line of best fit for the throughput to CPU
+    utilization ratio and shows more clearly the upward trend.
 
@@
 
-  %  throughput & CPU
-  \subsection{Throughput \& CPU}
-  These graphs show the interaction between CPU utilization and network throughput for each system with respect to IPFWD's acceptance probability.
-The dashed lines are CPU utilization, solid is network throughput in Mb/second.
-Higher acceptance probability means a greater chance that a given packet is immediately accepted without further processing.
-
-  \begin{center}
-  \includegraphics[width=0.45\textwidth]{cpu_isilon40}
-  \includegraphics[width=0.45\textwidth]{cpu_isilon10}
-  \includegraphics[width=0.45\textwidth]{cpu_wwu1}
-  \includegraphics[width=0.45\textwidth]{cpu_virtual10}
-  \end{center}
-
   % Difference in results between CPU and throughput bound systems
@@
 
-  All the systems tested can be broadly categorized into two groups: CPU bound
-  and NIC throughput bound. We can see that IPFWD affects performance
+    All the systems tested can be broadly categorized into two groups: CPU
+    bound and NIC throughput bound. We can see that IPFWD affects performance
     positively but differently for both types.
 
-  In CPU bound systems, IPFWD reduces the number of cycles it takes to process
-  each packet, allowing more packets to be processed in the same time and
-  increasing throughput. This is shown most clearly in the Virtualized system
-  tests, which have the weakest CPU but 10Gb Ethernet NICs. Regardless of test
-  type, the receiving system is always at 100\% CPU utilization. However, as
-  the probability to accept increases in IPFWD the throughput also increases.
-  This clearly shows that IPFWD decreases the CPU load required to process
-  incoming traffic.
+    In CPU bound systems, IPFWD reduces the number of cycles it takes to
+    process each packet, allowing more packets to be processed in the same time
+    and increasing throughput. This is shown most clearly in the Virtualised
+    system tests, which have the weakest CPU but 10Gb Ethernet NICs. Regardless
+    of test type, the receiving system is always at 100\% CPU utilization.
+    However, as the probability to accept increases in IPFWD the throughput
+    also increases.  This clearly shows that IPFWD decreases the CPU load
+    required to process incoming traffic.
 
-  In NIC throughput bound systems, reducing the number of CPU cycles to process
-  packets will not increase throughput. However, IPFWD still reduces CPU load
-  which allows more cycles to be dedicated to other system processes. This is
-  more difficult to see in the test results, particularly since the CPU
-  measurement tools are inherently imprecise and difficult to compare between
-  systems.
+    In NIC throughput bound systems, reducing the number of CPU cycles to
+    process packets will not increase throughput. However, IPFWD still reduces
+    CPU load which allows more cycles to be dedicated to other system
+    processes. This is more difficult to see in the test results, particularly
+    since the CPU measurement tools are inherently imprecise and difficult to
+    compare between systems.
+
+    % WWU 1G
+    However, the WWU 1G tests provide an example of a NIC bound system. The
+    lack of any performance gains in the WWU 1G tests provides an example of
+    when IPFWD should not be used. In this system, the CPU and other system
+    resources are more than sufficient to apply the entire IPFW ruleset to each
+    packet at wire-speed. Reducing the amount of CPU time it takes to process
+    each packet does not provide any benefit.
 
@@
 
-Firewall decision trees? (Complete Redundancy Detection)
-TODO
+  % Similiar work
+  Related work \cite{exploitingpacketsampling, analysisnetflow,
+  monitoringpacketsampling} into packet sampling in the space of network
+  security has been focused on detecting and classifying threats, not
+  performance enhancement.
 
-% Machine learning firewalls
+  %Firewall decision trees? (Complete Redundancy Detection)
 
 \section{Future Work}
+
   % Firewall rule reordering so we can "fail as early as possible"
-Ideally, a whitelist firewall's rule set would be constructed so that the most
-commonly applied matches are reached earlier rather than later. IPFWD
+  Ideally, a whitelist firewall's rule set would be constructed so that the
+  most commonly applied matches are reached earlier rather than later. IPFWD
   accomplishes this task by artificial early stopping. Firewall rule reordering
   is a could be another solution to this problem. Given logs of common network
-traffic for the system, statistical analysis could show which rules are matched
-more often than others. With this information, the rule set could be reordered
-so that the most commonly applied rules are earliest.
-TODO
+  traffic for the system, statistical analysis could show which rules are
+  matched more often than others. With this information, the rule set could be
+  reordered so that the most commonly applied rules are earliest.  TODO
+
+  % creating baseline invalid packet % so we can extrapolate even when we don't
+  % see any invalid packets
+
+  % have detection of bad traffic reduce acceptance rate
 
 \section{Conclusion}
-TODO
   % What performance benefit do we get?
-
   % What security cost is there?
-
   % Why is this useful?
-
   % Why is this important?
 
+  In performance sensitive computing environments, complex firewall rules
+  combined with high network throughput can result in heavy CPU load. IPFWD
+  addresses this problem by providing a graceful tradeoff between packet
+  visibility, total security policy enforcement and system performance.
+
+  IPFWD reduces the average amount of work to process packets by adding an
+  early rule to IPFD that has a chance to accept any packet. The security
+  implications of this behavior have been addressed and the performance
+  benefits provided have been demonstrated in a variety of systems.
+
+  IPFWD can be directly applied to high network performance and resource
+  conscious systems with complex firewall requirements to reduce CPU load and
+  increase network throughput.
+
+  This work also serves as a proof of concept for future probabilistic packet
+  matching firewall implementations.
+
 \appendices
-\section{Additional Figures}
-TODO
+\section{Additional Information}
+  TODO link to github repository
 
 % use section* for acknowledgement
-\section*{Acknowledgment}
-TODO
+%\section*{Acknowledgment}
+% TODO
 
@@
 % references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
-
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-% http://www.netperf.org/svn/netperf2/tags/netperf-2.7.0/doc/netperf.html#CPU-Utilization
-
-\end{thebibliography}
-
-% biography section
-% 
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
+\printbibliography
 
@@
 \end{document}
-
-% NOTES
-%==========================================================
-% needed in second column of first page if using \IEEEpubid
-%\IEEEpubidadjcol
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation Results}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command, the
-% \label for the overall figure must come after \caption.
-% \hfil must be used as a separator to get equal spacing.
-% The subfigure.sty package works much the same way, except \subfigure is
-% used instead of \subfloat.
-%
-%\begin{figure*}[!t]
-%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
-%\label{fig_second_case}}}
-%\caption{Simulation results}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-
-% An example of a floating table. Note that, for IEEE style tables, the 
-% \caption command should come BEFORE the table. Table text will default to
-% \footnotesize as IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that IEEE does not put floats in the very first column - or typically
-% anywhere on the first page for that matter. Also, in-text middle ("here")
-% positioning is not used. Most IEEE journals use top floats exclusively.
-% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
-% floats. This can be corrected via the \fnbelowfloat command of the
-% stfloats package.

### 2017-02-17 paper/main.tex
Rough draft complete

@@
   improved by reducing the number of rules applied to each packet. Supposing a
-  whitelist policy, a firewall has to apply every rule to each packet before
+  white-list policy, a firewall has to apply every rule to each packet before
   it's denied.  With IPFWD, the system has a chance to accept any packet early
@@
     policy by blocking packets that do not match the rules. The default action
-    for whitelist firewalls is to block all traffic, additional rules describe
+    for white-list firewalls is to block all traffic, additional rules describe
     exceptions. Blacklist firewalls are the opposite in that by default they
@@
     allowed. Since predicting and blacklisting all possible attacks is
-    infeasible, whitelist firewall rulesets are the accepted standard and the
+    infeasible, white-list firewall rule sets are the accepted standard and the
     focus of this paper.
@@
   and it's sufficient to be notified of security policy violations instead of
-  always stopping them \cite{TODO}.
+  always stopping them. These assumptions will not hold in all environments,
+  but we can leverage them for performance gains when they do.
 
@@
 
-  This is especially valuable for a whitelist firewall rule set dealing with
+  This is especially valuable for a white-list firewall rule set dealing with
   large amounts of rejected traffic; normally each rejected packet has all
@@
 
-  In a typical data center environment the vast majority of network traffic is
-  valid and going to be accepted eventually \cite{TODO}. However, these rule
-  sets are often complex and take increasing time to match a given packet
-  depending on the length. For a rule set of length $n$, we can expect to a
-  packet to encounter $\frac{n}{2}$ failed matches before matching the correct
-  rule and being accepted.
+  In a typical data center backend network the vast majority of traffic is
+  valid and going to be accepted by the firewall's rule set. However, these
+  rule sets are often complex and take increasing time to match packets
+  depending on the number of rules. For a rule set of length $n$, we can expect
+  to a packet to encounter $\frac{n}{2}$ failed matches before matching the
+  correct rule and being accepted.
 
@@
         IPFWD works by reducing the average number of rules that fail to match
-        before a packet is finished being processed. Very simple rulesets will
+        before a packet is finished being processed. Very simple rule sets will
         gain little from reducing the number of rules applied to each packet.
@@
   % Talk about security policy enforcement
-  Security policies define what is and is not allowed in an organization
-  \cite{TODO}. At the network and host level, this is implemented in part by a
-  firewall rule set. A policy violation on an internal network is more serious
-  than an external network. Internet facing hosts can expect to be scanned,
-  experience network anomalies and be attacked by third party agents
-  semi-regularly.
+  Security policies define what is and is not allowed in an organization . At
+  the network and host level, this is implemented in part by a firewall rule
+  set. A policy violation on an internal network is more serious than an
+  external network. Internet facing hosts can expect to be scanned, experience
+  network anomalies and be attacked by third party agents semi-regularly.
 
@@
       Malicious network activity and network errors are likely to be repeated
-      over time \cite{TODO}. The longer the activity persists, the more likely
-      it is to be detected by IPFW even when IPFWD is allowing a high
-      percentage of the network traffic through unchecked.
+      over time \cite{beacondetection}. The longer the activity persists, the
+      more likely it is to be detected by IPFW even when IPFWD is allowing a
+      high percentage of the network traffic through unchecked.
 
@@
     the target metric, which can be difficult to assess with TCP
-    \cite{TODO}.
+    \cite{netperf}.
 
@@
     when IPFWD should not be used. In this system, the CPU and other system
-    resources are more than sufficient to apply the entire IPFW ruleset to each
-    packet at wire-speed. Reducing the amount of CPU time it takes to process
-    each packet does not provide any benefit.
+    resources are more than sufficient to apply the entire IPFW rule set to
+    each packet at wire-speed. Reducing the amount of CPU time it takes to
+    process each packet does not provide any benefit.
 
 \section{Related Work}
-  % Cisco and Juniper routers
 
-  % Similiar work
-  Related work \cite{exploitingpacketsampling, analysisnetflow,
-  monitoringpacketsampling} into packet sampling in the space of network
-  security has been focused on detecting and classifying threats, not
-  performance enhancement.
+  % Packet sampling
+  Other papers \cite{exploitingpacketsampling, analysisnetflow,
+  monitoringpacketsampling} on packet sampling have been focused on detecting
+  and classifying threats, not performance enhancement. However, their
+  approaches to packet sampling and the implications addressed provided the
+  idea for IPFWD.
 
-  %Firewall decision trees? (Complete Redundancy Detection)
+  % IPFW performance
+  The BSD community has produced several in depth investigations into network
+  and firewall performance \cite{ipfwvspf,optimizingfreebsd}. Their work
+  informed the tests run to measure IPFWD's performance benefits.
 
@@
   % Firewall rule reordering so we can "fail as early as possible"
-  Ideally, a whitelist firewall's rule set would be constructed so that the
+  Ideally, a white-list firewall's rule set would be constructed so that the
   most commonly applied matches are reached earlier rather than later. IPFWD
@@
   matched more often than others. With this information, the rule set could be
-  reordered so that the most commonly applied rules are earliest.  TODO
+  reordered so that the most commonly applied rules are earliest.
+
+  Other work \cite{redundancydetection} has been done in removing redundant
+  rules to increase firewall performance. It's possible that a combination of
+  redundancy checking and priority reordering could provide further performance
+  gains.
 
@@
   % see any invalid packets 
+  Machine learning could be applied to packet samples to more accurately
+  determine appropriate bounds on the early acceptance probability. Different
+  systems and circumstances may require more or less stringent packet checking.
+  Environments with very high confidence in network security might raise the
+  lower bound of the acceptance rate to improve performance. Likewise, less
+  confident environments may lower the upper bound to ensure an acceptable
+  percentage of packets are always checked.
 
   % have detection of bad traffic reduce acceptance rate
+  Future versions of IPFWD could benefit from additional modifiers to the early
+  acceptance probability. For instance, the system could reduce the early
+  acceptance probability whenever packets are dropped by the firewall under the
+  assumption that invalid traffic likely precedes more invalid traffic.
 
@@
   combined with high network throughput can result in heavy CPU load. IPFWD
-  addresses this problem by providing a graceful tradeoff between packet
+  addresses this problem by providing a graceful trade off between packet
   visibility, total security policy enforcement and system performance.
@@
   IPFWD reduces the average amount of work to process packets by adding an
-  early rule to IPFD that has a chance to accept any packet. The security
+  early rule to IPFW that has a chance to accept any packet. The security
   implications of this behavior have been addressed and the performance
@@
   conscious systems with complex firewall requirements to reduce CPU load and
-  increase network throughput.
-
-  This work also serves as a proof of concept for future probabilistic packet
-  matching firewall implementations.
+  increase network throughput. This work also serves as a proof of concept for
+  future probabilistic packet matching firewall implementations.
 
@@
 \section{Additional Information}
-  TODO link to github repository
+  The source code for IPFWD, additional test results, and development cycle
+  information may be found at \texttt{https://???}.
 

### 2017-02-22 paper/_main.tex
Huge clean up for final paper, paper changes

--- 2017-02-17 paper/main.tex
@@
 
+\usepackage{caption}
 \usepackage{graphicx}
@@
 
-\author{Philip Nelson and Austin Voecks% <-this % stops a space
+\author{Austin Voecks and Philip Nelson% <-this % stops a space
 \thanks{Supported in part by Dell EMC Isilon through grant and hardware}}
@@
 lot of packet processing. This can negatively impact performance-sensitive
-systems such as storage nodes. This paper describes a practical solution to
-this problem using a load-weighted probabilistic mechanism that allows a
-trade-off between perfect visibility of incoming packets and reduced impact to
-system load.
+systems such as storage nodes in data centers. This paper describes a
+practical solution to this problem using a load-weighted probabilistic
+mechanism that allows a trade-off between perfect visibility of incoming
+packets and reduced impact to system load.
 
@@
 
-% TODO rewrite
 \section{Introduction}
 
-  IPFWD is a daemon for FreeBSD that updates an early rule in IPFW that has a
-  chance to accept any packet. The probability of acceptance is updated over
-  time and dependent on the current system load. In high performance systems
-  this results in an increase in network throughput.
+  IPFWD is a daemon for FreeBSD that updates an early rule in the IP Firewall
+  (IPFW) that has a chance to accept any packet. The probability of early
+  acceptance is updated over time and dependent on the current system load. In
+  high performance systems this can result in an increase in network
+  throughput.
 
@@
   improved by reducing the number of rules applied to each packet. Supposing a
-  white-list policy, a firewall has to apply every rule to each packet before
+  white-list policy, a firewall must apply every rule to a packet before
   it's denied.  With IPFWD, the system has a chance to accept any packet early
-  and skip any further computation. Test results shows this reduces the system
-  resources required to handle the same amount of traffic.
+  and skip any further computation. Test results show this reduces the 
+  resources required to handle the same amount of traffic in some systems.
 
   This is a shift in mindset from typical firewalls \cite{networksecurity}.
-  Instead of enforcing every part of a security policy all the time, IPFWD
-  enforces the policy some of the time and extrapolates from the violations it
-  encounters.  This provides insight into the violations that weren't detected.
-  For this cost, you gain increased firewall performance and network throughput
-  in resource bound systems. It's acceptable to allow a percentage of policy
+  Instead of enforcing every part of the security policy all the time, IPFWD
+  enforces the policy some of the time provides additional information so the
+  missed violations can be inferred. For this cost, you gain increased firewall
+  performance and network throughput in resource bound systems. IPFWD works
+  under the assumption that it's acceptable to allow a percentage of policy
   violations given that network traffic patterns are often repeated and the
-  goal is detection and not immediate prevention. Preventative action may be
+  goal is detection, not immediate prevention. Administrative action may be
   taken later when resource requirements are lower.
@@
   As an example, under heavy load IPFWD may immediately accept 40\% of packets.
-  Supposing a port scan was initiated during this time and rules exist to block
-  it, 60\% of the port scan would still be rejected and logged. Since the
+  Some of those packets may have been malicious. Supposing a port scan was
+  initiated during this time and rules exist to block it, at least 60\% of the
+  port scan would still be rejected and logged. Since the early acceptance
   probability will fluctuate over time, IPFWD provides information in the IPFW
@@
     according to a security policy. Without a security policy, administrators
-    cannot know what to allow and disallow. Typical firewalls enforce the
-    policy by blocking packets that do not match the rules. The default action
-    for white-list firewalls is to block all traffic, additional rules describe
-    exceptions. Blacklist firewalls are the opposite in that by default they
-    accept all traffic and additional rules describe packet types that are not
-    allowed. Since predicting and blacklisting all possible attacks is
-    infeasible, white-list firewall rule sets are the accepted standard and the
-    focus of this paper.
+    cannot know what to allow and disallow when constructing firewall rules.
+    Typical firewalls enforce the policy by blocking packets that do not match
+    any rules. The default action for white-list firewalls is to block all
+    traffic, additional rules describe exceptions. Blacklist firewalls are the
+    opposite, by default they accept all traffic and additional rules describe
+    packet types that are not allowed. Since predicting and blacklisting all
+    possible attacks is not possible in practice, white-list firewall rule sets
+    are the accepted standard and the focus of this paper.
 
@@
     and IPv6. It is comprised of several components: a kernel firewall filter
-    rule processor and its integrated packet accounting facility, the logging
+    rule processor and its integrated packet accounting facility, a logging
     facility, NAT, the dummynet traffic shaper, a forward facility, a bridge
@@
 
-    IPFW was chosen for it's close integration with the FreeBSD operating
-    system, in kernel packet filter, and general performance. Performance is
-    dependent on system parameters and circumstances, but IPFW outperforms PF
-    in stateful packet filtering \cite{ipfwvspf}. Stateful rules are more
-    powerful and flexible than stateless rules, were no session information is
-    maintained \cite{networksecurity}.  Additionally, the capabilities of IPFW
-    extend beyond packet filtering into source based routing, traffic shaping
-    and more.
+    IPFW was chosen for its close integration with the FreeBSD operating
+    system, kernel packet filter, and performance. Performance is dependent on
+    system parameters and circumstances, but IPFW outperforms PF in stateful
+    packet filtering \cite{ipfwvspf}. Stateful rules are more powerful and
+    flexible than stateless rules, were no session information is maintained
+    \cite{networksecurity}.  Additionally, the capabilities of IPFW extend
+    beyond packet filtering into source based routing, traffic shaping and
+    more. It's for these reasons that IFPW was chosen over PF.
 
@@
     matching. From the documentation, this feature was intended for load
-    balancing and other traffic shaping tasks. It's the core of IPFWD's
+    balancing and other traffic shaping tasks. It's the core of IPFWDs
     operation.
@@
   It does this by adding and maintaining an early rule in the IPFW rule set of
-  the following form: \verb|prob 0.000 allow ip from any to any|.
+  the following form: 
+  \begin{center}
+      \verb|prob 0.000 allow ip from any to any|
+  \end{center}
+  which allows any IP packet, incoming or outgoing with a chance equal to the probability given.
 
@@
   % - Probabilistic packet matching
-  IPFWD's behavior is based on the assumption that the majority of network
+  IPFWDs behavior is based on the assumption that the majority of network
   traffic is valid, malicious events are rare \cite{networktrafficanalysis},
@@
   encountered before the majority of the other rules in the rule set.  The
-  probability given in this rule determines the chance that we skip processing
-  all further rules, supposing that there hasn't been a match already.
+  probability given in this rule determines the chance that IPFWD skips
+  processing all further rules, supposing that there has not been a match
+  already.
 
@@
   large amounts of rejected traffic; normally each rejected packet has all
-  rules applied to it while IPFW looks for a match.  When no matches are
+  rules applied to it in order to find a match. When no matches are
   encountered, the default rule is employed and the packet is rejected.
 
-  In a typical data center backend network the vast majority of traffic is
-  valid and going to be accepted by the firewall's rule set. However, these
+  In a typical data center backend network, the vast majority of traffic is
+  valid and going to be accepted by the firewalls rule set. However, these
   rule sets are often complex and take increasing time to match packets
   depending on the number of rules. For a rule set of length $n$, we can expect
-  to a packet to encounter $\frac{n}{2}$ failed matches before matching the
-  correct rule and being accepted.
+  a packet to encounter on average $\frac{n}{2}$ failed matches before matching
+  the correct rule and being accepted.
 
@@
   creates a chance that packets that would have normally been rejected will get
-  through the firewall. We'll refer to these types of packets as invalid for
-  the purpose of discussion.
+  through the firewall.
 
@@
         reducing processing time. Given modern CPU clock speeds, systems with
-        network cards slower than 1 Gb/sec will likely not benefit either.
+        network cards slower than 1 Gb/sec will likely not benefit from IPFWD.
 
@@
         information can be used to infer the existence of additional invalid
-        packets.
+        packets. Environments that cannot allow this relaxation in security
+        policy enforcement will no benefit from IPFWD.
 
@@
   % Talk about security policy enforcement
-  Security policies define what is and is not allowed in an organization . At
+  Security policies define what is and is not allowed in an organization. At
   the network and host level, this is implemented in part by a firewall rule
@@
   external network. Internet facing hosts can expect to be scanned, experience
-  network anomalies and be attacked by third party agents semi-regularly.
+  network anomalies, and be attacked by third party agents more often than
+  internal hosts.
 
-  However when this happens on an internal network, there is a strong
-  possibility something else is going wrong. The usage of IPFWD is predicated
-  on the assumption that it's sufficient to be notified of these events so
-  further action can be taken later than to always immediately stop them.
+  However, when these events do occur on an internal network, there is a strong
+  possibility something else is wrong. Unexpected network activity on a fully
+  controlled internal network indicates compromise or misconfiguration, both of
+  which are serious problems. The usage of IPFWD is predicated on the
+  assumption that it's sufficient to be notified of these events so further
+  action can be taken later.
 
@@
   inspection does mean that attacks that would have normally been blocked could
-  succeed. In addition to performance improvements, the following factors help
-  to account for this risk:
+  succeed. The following factors help to account for this risk:
 
@@
 
-      Malicious network activity and network errors are likely to be repeated
-      over time \cite{beacondetection}. The longer the activity persists, the
-      more likely it is to be detected by IPFW even when IPFWD is allowing a
-      high percentage of the network traffic through unchecked.
+      Malicious network activity such as malware beaconing and network
+      reconnaissance, as well as general network errors are likely to be
+      repeated over time \cite{beacondetection}. The longer the activity
+      persists, the more likely it is to be detected by IPFW, even when IPFWD
+      is allowing a high percentage of the network traffic through unchecked.
 
@@
       per-packet basis, malicious network activity will experience
-      approximately $m\%$ packet loss where $m$ is IPFWD's probability of early
+      approximately $(1 - m)\%$ packet loss where $m$ is IPFWDs probability of early
       acceptance. The exact effect this would have on malware or reconnaissance
-      would depend, but generally will increase the chance they fail or produce
+      would depend, but generally will increase the chance of failure or
       invalid results.
@@
   % How do you trust this system?
-  Knowledge of a network's traffic patterns and the additional log information
-  provided by IPFWD can help. It allows administrators to extrapolate the
-  existence of undetected network events through simple statistical models.
 
@@
   % - Is this really different than traditional firewalls?
-  TODO
 
@@
 
-    Suppose each host in a data center's internal network is running IPFWD on
-    each host, average CPU usage is $75\%$, and have rule sets blocking common
-    network attacks. One of the hosts has been compromised and begins
-    conducting network reconnaissance by port scanning the other hosts on the
-    network.
+    Suppose each host in a data centers internal network is running IPFW and
+    IPFWD on each host, average CPU usage is $75\%$, and have rule sets
+    blocking common network attacks. One of the hosts has been compromised and
+    begins conducting network reconnaissance by port scanning the other hosts
+    on the network.
 
-    Firewall rules are in place to block host to host communication $m\%$ of
+    Firewall rules are in place to block host-to-host communication to $m\%$ of
     the ports being scanned. This means that $1 - m\%$ of the scan would be
-    allowed regardless of IPFWD's action. This percentage represents valid
+    allowed regardless of IPFWDs action. This percentage represents valid
     host-to-host communication as determined by the security policy.
@@
   % describe difference between reply and no reply violations
-  \subsection{Security Cost}
-
-    The chance of missing a security policy violation depends on the type of
-    traffic. The following situations exist when both incoming and outgoing
+  Thus, the chance of missing a security policy violation depends on the type
+  of traffic. The following situations exist when both incoming and outgoing
   firewall rules are employed:
@@
       Here, the firewall only has one chance to catch the policy violation.
-        So if the early acceptance probability is $m$, then $m\%$ of the policy
+      So, if the early acceptance probability is $m$, then $m\%$ of the policy
       violating packets will be allowed.
@@
 
-        Here, the firewall has two chances to catch the policy violation.  So
+      Here, the firewall has two chances to catch the policy violation. So,
       if the early acceptance probability is $m$, then $m'\%$ of the policy
@@
 
-  Tests were run on three platforms and five different network cards, resulting
-  in five different systems. The most important parameters to our tests are
-  CPU, NIC, and ENV. Descriptions of each system are given in the tables
+  Tests were run on three platforms and four different network cards, referred
+  to as four different systems. The most important parameters to these tests
+  are CPU, NIC, and ENV. Descriptions of each system are given in the tables
   below:\\
@@
 
-  Virtualisation makes it more difficult to have the same confidence in the
+  Virtualisation made it more difficult to have the same confidence in the
   results as with the physical machines. Tests on the Virtualised system were
@@
   The WWU and Virtualised tests were conducted on stock FreeBSD 11 without
-  additional network performance configuration applied. Isilon OneFS does
-  contain additional network performance tuning and demonstrates the additional
-  gains IPFWD can provide with careful FreeBSD 11 kernel configuration.
+  additional network performance configuration applied. The Isilon systems
+  contained additional network performance tuning and demonstrates the
+  additional gains IPFWD can provide with careful FreeBSD 11 kernel
+  configuration.
 
@@
     compiled using Netperf. Netperf was chosen because it was readily available
-    on all test platforms, provides built-in CPU utilization measurement, and a
-    wide range of test types \cite{netperf}.
+    on all test platforms, provides built-in CPU utilization measurement, has a
+    wide range of test types \cite{netperf}, and is used elsewhere in the BSD
+    community for testing.
 
@@
     The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to
-    preserve CPU cycles for IPFW. Raw network throughput and CPU performance is
-    the target metric, which can be difficult to assess with TCP
+    preserve CPU cycles for IPFW. Raw network throughput and CPU performance
+    was the target metric, which can be difficult to assess with TCP
     \cite{netperf}.
 
-    Netperf is single threaded, which allows easier comparison of results
+    Netperf is single threaded, which allowed easier comparison of results
     between systems with different numbers of cores. Additionally, all tests
@@
 
-    CPU utilization measurement is the most difficult parameter to measure in
+    CPU utilization measurement was the most difficult parameter to measure in
     these tests. There are many factors that can account for variability
     between tests and jitter in the results. The Netperf manual describes these
-    challenges and the variety of techniques used on different systems.
-    \cite{netperf}
+    challenges and the variety of techniques used on different systems to
+    measure CPU usage \cite{netperf}.
 
-    Netperf was run in single threaded mode, which explains why throughput
-    wasn't closer to the maximum throughput for the NIC on each system. The
+    Netperf was run in single threaded mode, which explains why throughput was
+    not closer to the maximum throughput for the NIC on each system. The
     FreeBSD network stack allows multithreaded packet processing by utilizing
-    multiple queues \cite{freebsdhandbook}. Since we're indirectly testing the
-    performance of IPFW through Netperf, there are two different CPU
+    multiple queues meaning the kernel can spread the work done by IPFW over
+    multiple cores \cite{freebsdhandbook}. Since we were indirectly testing the
+    performance of IPFW through Netperf, there were two different CPU
     measurements to account for: kernel time and user time. Since IPFW runs in
-    the kernel, it's CPU time is classified as kernel time.
+    the kernel, its CPU time is classified as kernel time.
 
-    Netperf was able to measure CPU utilization on the Virtualized and WWU
+    Netperf was able to measure CPU utilization on the Virtualised and WWU
     systems, but didn't work on the Isilon systems. Instead, \textit{ps} was
     used. These tests are comparable since both measured user CPU utilization,
-    not kernel CPU utilization.
+    not kernel CPU utilization. Thus, IPFWs CPU usage was measured indirectly
+    through the CPU usage of netperf.
 
@@
     CPU utilization is high and throughput is lower than the NIC maximum (eg:
-    Virtualised system). Conversely, when CPU availablility exceeds NIC
-    throughput, throughput is high and CPU utilization is low (eg: WWU 1G
+    the Virtualised system). Conversely, when CPU availability exceeds NIC
+    throughput, throughput is high and CPU utilization is low (eg: the WWU 1G
     system).
@@
 
-    These graphs show the interaction between CPU utilization and network
-    throughput for each system with respect to IPFWD's early acceptance
-    probability. The dashed lines are CPU utilization, solid is network
-    throughput in Mb/second. Higher acceptance probability means a greater
-    chance that a given packet is immediately accepted without further
+    The graphs in figure \ref{fig:isiloncpu40}, \ref{fig:isiloncpu10}, \ref{fig:wwucpu1}, and \ref{fig:virtualcpu10} show the interaction between CPU utilization and network
+    throughput for each system with respect to IPFWDs early acceptance
+    probability. The dashed lines are CPU utilization; solid is network
+    throughput in Mb/second. A higher acceptance probability means a greater
+    chance that a given packet was immediately accepted without further
     processing.
@@
     acceptance probability increases, so does throughput. The CPU usage shows
-    that this throughput is gained without additional CPU utilization, meaning
-    the system is spending fewer cycles to process each packet on average.
+    that this additional throughput was gained without additional CPU
+    utilization, meaning the system was spending fewer cycles to process each
+    packet on average.
 
-    The lack of any trend in the WWU 1G graph is expected and addressed in the
+    The lack of any trend in the WWU 1G graph was expected and addressed in the
     \textit{Discussion} section.
 
-    \begin{center}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{cpu_isilon40}
+      \caption{Throughput \& CPU for Isilon 40G system}
+      \label{fig:isiloncpu40}
+    \end{figure}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{cpu_isilon10}
+      \caption{Throughput \& CPU for Isilon 10G system}
+      \label{fig:isiloncpu10}
+    \end{figure}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{cpu_wwu1}
+      \caption{Throughput \& CPU for WWU 1G system}
+      \label{fig:wwucpu1}
+    \end{figure}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{cpu_virtual10}
-    \end{center}
+      \caption{Throughput \& CPU for Virtualised 10G system}
+      \label{fig:virtualcpu10}
+    \end{figure}
+
+    %\begin{center}
+    %  \includegraphics[width=0.45\textwidth]{cpu_isilon40}
+    %  \includegraphics[width=0.45\textwidth]{cpu_isilon10}
+    %  \includegraphics[width=0.45\textwidth]{cpu_wwu1}
+    %  \includegraphics[width=0.45\textwidth]{cpu_virtual10}
+    %\end{center}
 
@@
 
-    These graphs show network the ratio of throughput to CPU utilization for
+    The graphs in figures \ref{fig:isilontoc40}, \ref{fig:isilontoc10}, \ref{fig:wwutoc1}, and \ref{fig:virtualtoc10}  show network the ratio of throughput to CPU utilization for
     each system where $ratio(x) = throughput(x) / cpu(x)$. The data being
-    presented here is the same as the previous graphs, shown differently. The
-    upward trend over the x-axis shows that as the probability of early
-    acceptance increases, the system is spending less CPU time to process
-    packets on average.
+    presented here is the same as the previous graphs, merely shown
+    differently. The upward trend over the x-axis shows that as the probability
+    of early acceptance increases, the system is spending less CPU time to
+    process packets on average.
 
     The dashed line shows the line of best fit for the throughput to CPU
-    utilization ratio and shows more clearly the upward trend.
+    utilization ratio and shows the upward trend more clearly than the raw
+    data.
 
-    \begin{center}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{toc_isilon40}
+      \caption{Throughput/CPU for Isilon 40G system}
+      \label{fig:isilontoc40}
+    \end{figure}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{toc_isilon10}
+      \caption{Throughput/CPU for Isilon 10G system}
+      \label{fig:isilontoc10}
+    \end{figure}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{toc_wwu1}
+      \caption{Throughput/CPU for WWU 1G system}
+      \label{fig:wwutoc1}
+    \end{figure}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{toc_virtual10}
-    \end{center}
+      \caption{Throughput/CPU for Virtualised 10G system}
+      \label{fig:virtualtoc10}
+    \end{figure}
+
+    %\begin{center}
+    %  \includegraphics[width=0.45\textwidth]{toc_isilon40}
+    %  \includegraphics[width=0.45\textwidth]{toc_isilon10}
+    %  \includegraphics[width=0.45\textwidth]{toc_wwu1}
+    %  \includegraphics[width=0.45\textwidth]{toc_virtual10}
+    %\end{center}
 
@@
     However, as the probability to accept increases in IPFWD the throughput
-    also increases.  This clearly shows that IPFWD decreases the CPU load
+    also increases.  This clearly shows that IPFWD decreases the CPU time
     required to process incoming traffic.
@@
   approaches to packet sampling and the implications addressed provided the
-  idea for IPFWD.
+  inspiration for IPFWD.
 
@@
   and firewall performance \cite{ipfwvspf,optimizingfreebsd}. Their work
-  informed the tests run to measure IPFWD's performance benefits.
+  informed the tests run here.
 
@@
   % Firewall rule reordering so we can "fail as early as possible"
-  Ideally, a white-list firewall's rule set would be constructed so that the
-  most commonly applied matches are reached earlier rather than later. IPFWD
+  Ideally, a white-list firewalls rule set would be constructed so that the
+  most commonly applied matches are reached sooner rather than later. IPFWD
   accomplishes this task by artificial early stopping. Firewall rule reordering
-  is a could be another solution to this problem. Given logs of common network
+  could be another solution to this problem. Given logs of common network
   traffic for the system, statistical analysis could show which rules are
@@
   Other work \cite{redundancydetection} has been done in removing redundant
-  rules to increase firewall performance. It's possible that a combination of
+  rules to increase firewall performance. It is possible that a combination of
   redundancy checking and priority reordering could provide further performance
@@
   determine appropriate bounds on the early acceptance probability. Different
-  systems and circumstances may require more or less stringent packet checking.
+  systems and environments may require more or less stringent packet checking.
   Environments with very high confidence in network security might raise the
@@
   acceptance probability. For instance, the system could reduce the early
-  acceptance probability whenever packets are dropped by the firewall under the
-  assumption that invalid traffic likely precedes more invalid traffic.
+  acceptance probability whenever packets are dropped by the firewall, under
+  the assumption that invalid traffic likely precedes more invalid traffic.
 
@@
   In performance sensitive computing environments, complex firewall rules
-  combined with high network throughput can result in heavy CPU load. IPFWD
-  addresses this problem by providing a graceful trade off between packet
-  visibility, total security policy enforcement and system performance.
+  combined with high network throughput can result in heavy CPU load, limiting
+  system performance. IPFWD addresses this problem by providing a graceful
+  trade off between packet visibility, total security policy enforcement and
+  system performance.
 
@@
   The source code for IPFWD, additional test results, and development cycle
-  information may be found at \texttt{https://???}.
-
-% use section* for acknowledgement
-%\section*{Acknowledgment}
-% TODO
+  information may be found at \texttt{https://TODO}.
 

### 2017-02-22 downloads/top/commands.c

@@ -1,541 +1,0 @@
-/*
- *  Top users/processes display for Unix
- *  Version 3
- *
- *  This program may be freely redistributed,
- *  but this entire comment MUST remain intact.
- *
- *  Copyright (c) 1984, 1989, William LeFebvre, Rice University
- *  Copyright (c) 1989, 1990, 1992, William LeFebvre, Northwestern University
- *
- * $FreeBSD: releng/11.0/contrib/top/commands.c 300395 2016-05-22 04:17:00Z ngie $
- */
-
-/*
- *  This file contains the routines that implement some of the interactive
- *  mode commands.  Note that some of the commands are implemented in-line
- *  in "main".  This is necessary because they change the global state of
- *  "top" (i.e.:  changing the number of processes to display).
- */
-
-#include "os.h"
-
-#include <sys/time.h>
-#include <sys/resource.h>
-
-#include <ctype.h>
-#include <errno.h>
-#include <signal.h>
-#include <unistd.h>
-
-#include "commands.h"
-#include "sigdesc.h"		/* generated automatically */
-#include "top.h"
-#include "boolean.h"
-#include "utils.h"
-#include "machine.h"
-
-extern int  errno;
-
-extern char *copyright;
-
-/* imported from screen.c */
-extern int overstrike;
-
-int err_compar();
-char *err_string();
-static int str_adderr(char *str, int len, int err);
-static int str_addarg(char *str, int len, char *arg, int first);
-
-/*
- *  show_help() - display the help screen; invoked in response to
- *		either 'h' or '?'.
- */
-
-void
-show_help()
-
-{
-    printf("Top version %s, %s\n", version_string(), copyright);
-    fputs("\n\n\
-A top users display for Unix\n\
-\n\
-These single-character commands are available:\n\
-\n\
-^L      - redraw screen\n\
-q       - quit\n\
-h or ?  - help; show this text\n", stdout);
-
-    /* not all commands are availalbe with overstrike terminals */
-    if (overstrike)
-    {
-	fputs("\n\
-Other commands are also available, but this terminal is not\n\
-sophisticated enough to handle those commands gracefully.\n\n", stdout);
-    }
-    else
-    {
-	fputs("\
-C       - toggle the displaying of weighted CPU percentage\n\
-d       - change number of displays to show\n\
[…]

### 2017-02-22 downloads/top/commands.h

@@ -1,21 +1,0 @@
-/*
- *  Top users/processes display for Unix
- *  Version 3
- *
- *  This program may be freely redistributed,
- *  but this entire comment MUST remain intact.
- *
- *  Copyright (c) 1984, 1989, William LeFebvre, Rice University
- *  Copyright (c) 1989, 1990, 1992, William LeFebvre, Northwestern University
- *  Copyright (c) 2016, Randy Westlund
- *
- * $FreeBSD: releng/11.0/contrib/top/commands.h 300395 2016-05-22 04:17:00Z ngie $
- */
-#ifndef COMMANDS_H
-#define COMMANDS_H
-
-void	show_errors(void);
-int	error_count(void);
-void	show_help(void);
-
-#endif /* COMMANDS_H */

### 2017-02-22 downloads/top/display.c

@@ -1,1399 +1,0 @@
-/*
- *  Top users/processes display for Unix
- *  Version 3
- *
- *  This program may be freely redistributed,
- *  but this entire comment MUST remain intact.
- *
- *  Copyright (c) 1984, 1989, William LeFebvre, Rice University
- *  Copyright (c) 1989, 1990, 1992, William LeFebvre, Northwestern University
- *
- * $FreeBSD: releng/11.0/contrib/top/display.c 300395 2016-05-22 04:17:00Z ngie $
- */
-
-/*
- *  This file contains the routines that display information on the screen.
- *  Each section of the screen has two routines:  one for initially writing
- *  all constant and dynamic text, and one for only updating the text that
- *  changes.  The prefix "i_" is used on all the "initial" routines and the
- *  prefix "u_" is used for all the "updating" routines.
- *
- *  ASSUMPTIONS:
- *        None of the "i_" routines use any of the termcap capabilities.
- *        In this way, those routines can be safely used on terminals that
- *        have minimal (or nonexistant) terminal capabilities.
- *
- *        The routines are called in this order:  *_loadave, i_timeofday,
- *        *_procstates, *_cpustates, *_memory, *_message, *_header,
- *        *_process, u_endscreen.
- */
-
-#include "os.h"
-
-#include <sys/time.h>
-
-#include <ctype.h>
-#include <time.h>
-#include <unistd.h>
-
-#include "screen.h"		/* interface to screen package */
-#include "layout.h"		/* defines for screen position layout */
-#include "display.h"
-#include "top.h"
-#include "top.local.h"
-#include "boolean.h"
-#include "machine.h"		/* we should eliminate this!!! */
-#include "utils.h"
-
-#ifdef DEBUG
-FILE *debug;
-#endif
-
-/* imported from screen.c */
-extern int overstrike;
-
-static int lmpid = 0;
-static int last_hi = 0;		/* used in u_process and u_endscreen */
-static int lastline = 0;
-static int display_width = MAX_COLS;
-
-#define lineindex(l) ((l)*display_width)
-
-
-/* things initialized by display_init and used thruout */
-
-/* buffer of proc information lines for display updating */
-char *screenbuf = NULL;
-
-static char **procstate_names;
[…]

### 2017-02-22 downloads/top/layout.h

@@ -1,33 +1,0 @@
-/*
- *  Top - a top users display for Berkeley Unix
- *
- *  This file defines the locations on tne screen for various parts of the
- *  display.  These definitions are used by the routines in "display.c" for
- *  cursor addressing.
- *
- * $FreeBSD: releng/11.0/contrib/top/layout.h 237656 2012-06-27 18:08:48Z jhb $
- */
-
-extern int  x_lastpid;		/* 10 */
-extern int  y_lastpid;		/* 0 */
-extern int  x_loadave;		/* 33 */
-extern int  x_loadave_nompid;	/* 15 */
-extern int  y_loadave;		/* 0 */
-extern int  x_procstate;	/* 0 */
-extern int  y_procstate;	/* 1 */
-extern int  x_brkdn;		/* 15 */
-extern int  y_brkdn;		/* 1 */
-extern int  x_mem;		/* 5 */
-extern int  y_mem;		/* 3 */
-extern int  x_arc;		/* 5 */
-extern int  y_arc;		/* 4 */
-extern int  x_swap;		/* 6 */
-extern int  y_swap;		/* 4 */
-extern int  y_message;		/* 5 */
-extern int  x_header;		/* 0 */
-extern int  y_header;		/* 6 */
-extern int  x_idlecursor;	/* 0 */
-extern int  y_idlecursor;	/* 5 */
-extern int  y_procs;		/* 7 */
-
-extern int  y_cpustates;	/* 2 */

### 2017-02-22 downloads/top/machine.c

@@ -1,1649 +1,0 @@
-/*
- * top - a top users display for Unix
- *
- * SYNOPSIS:  For FreeBSD-2.x and later
- *
- * DESCRIPTION:
- * Originally written for BSD4.4 system by Christos Zoulas.
- * Ported to FreeBSD 2.x by Steven Wallace && Wolfram Schneider
- * Order support hacked in from top-3.5beta6/machine/m_aix41.c
- *   by Monte Mitzelfelt (for latest top see http://www.groupsys.com/topinfo/)
- *
- * This is the machine-dependent module for FreeBSD 2.2
- * Works for:
- *	FreeBSD 2.2.x, 3.x, 4.x, and probably FreeBSD 2.1.x
- *
- * LIBS: -lkvm
- *
- * AUTHOR:  Christos Zoulas <christos@ee.cornell.edu>
- *          Steven Wallace  <swallace@freebsd.org>
- *          Wolfram Schneider <wosch@FreeBSD.org>
- *          Thomas Moestl <tmoestl@gmx.net>
- *
- * $FreeBSD$
- */
-
-#include <sys/param.h>
-#include <sys/errno.h>
-#include <sys/file.h>
-#include <sys/proc.h>
-#include <sys/resource.h>
-#include <sys/rtprio.h>
-#include <sys/signal.h>
-#include <sys/sysctl.h>
-#include <sys/time.h>
-#include <sys/user.h>
-#include <sys/vmmeter.h>
-
-#include <err.h>
-#include <kvm.h>
-#include <math.h>
-#include <nlist.h>
-#include <paths.h>
-#include <pwd.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <strings.h>
-#include <unistd.h>
-#include <vis.h>
-
-#include "top.h"
-#include "machine.h"
-#include "screen.h"
-#include "utils.h"
-#include "layout.h"
-
-#define GETSYSCTL(name, var) getsysctl(name, &(var), sizeof(var))
-#define	SMPUNAMELEN	13
-#define	UPUNAMELEN	15
-
-extern struct process_select ps;
-extern char* printable(char *);
-static int smpmode;
-enum displaymodes displaymode;
-#ifdef TOP_USERNAME_LEN
-static int namelength = TOP_USERNAME_LEN;
-#else
-static int namelength = 8;
-#endif
-/* TOP_JID_LEN based on max of 999999 */
-#define TOP_JID_LEN 7
-#define TOP_SWAP_LEN 6
[…]

### 2017-02-22 downloads/top/machine.h

@@ -1,91 +1,0 @@
-/*
- * $FreeBSD: releng/11.0/contrib/top/machine.h 300395 2016-05-22 04:17:00Z ngie $
- */
-
-/*
- *  This file defines the interface between top and the machine-dependent
- *  module.  It is NOT machine dependent and should not need to be changed
- *  for any specific machine.
- */
-#ifndef MACHINE_H
-#define MACHINE_H
-
-#include "top.h"
-
-/*
- * the statics struct is filled in by machine_init
- */
-struct statics
-{
-    char **procstate_names;
-    char **cpustate_names;
-    char **memory_names;
-    char **arc_names;
-    char **swap_names;
-#ifdef ORDER
-    char **order_names;
-#endif
-    int ncpus;
-};
-
-/*
- * the system_info struct is filled in by a machine dependent routine.
- */
-
-#ifdef p_active     /* uw7 define macro p_active */
-#define P_ACTIVE p_pactive
-#else
-#define P_ACTIVE p_active
-#endif
-
-struct system_info
-{
-    int    last_pid;
-    double load_avg[NUM_AVERAGES];
-    int    p_total;
-    int    P_ACTIVE;     /* number of procs considered "active" */
-    int    *procstates;
-    int    *cpustates;
-    int    *memory;
-    int    *arc;
-    int    *swap;
-    struct timeval boottime;
-    int    ncpus;
-};
-
-/* cpu_states is an array of percentages * 10.  For example, 
-   the (integer) value 105 is 10.5% (or .105).
- */
-
-/*
- * the process_select struct tells get_process_info what processes we
- * are interested in seeing
- */
-
-struct process_select
-{
-    int idle;		/* show idle processes */
-    int self;		/* show self */
-    int system;		/* show system processes */
-    int thread;		/* show threads */
-    int uid;		/* only this uid (unless uid == -1) */
-    int wcpu;		/* show weighted cpu */
-    int jid;		/* only this jid (unless jid == -1) */
-    int jail;		/* show jail ID */
-    int kidle;		/* show per-CPU idle threads */
-    char *command;	/* only this command (unless == NULL) */
-};
-
-/* routines defined by the machine dependent module */
-
-char	*format_header();
-char	*format_next_process();
-void	 toggle_pcpustats(void);
-void	 get_system_info(struct system_info *si);
-int		 machine_init(struct statics *statics, char do_unames);
-int		 proc_owner(int pid);
-
[…]

### 2017-02-22 downloads/top/screen.c

@@ -1,507 +1,0 @@
-/*
- *  Top users/processes display for Unix
- *  Version 3
- *
- *  This program may be freely redistributed,
- *  but this entire comment MUST remain intact.
- *
- *  Copyright (c) 1984, 1989, William LeFebvre, Rice University
- *  Copyright (c) 1989, 1990, 1992, William LeFebvre, Northwestern University
- *
- * $FreeBSD: releng/11.0/contrib/top/screen.c 300395 2016-05-22 04:17:00Z ngie $
- */
-
-/*  This file contains the routines that interface to termcap and stty/gtty.
- *
- *  Paul Vixie, February 1987: converted to use ioctl() instead of stty/gtty.
- *
- *  I put in code to turn on the TOSTOP bit while top was running, but I
- *  didn't really like the results.  If you desire it, turn on the
- *  preprocessor variable "TOStop".   --wnl
- */
-
-#include "os.h"
-#include "top.h"
-
-#include <sys/ioctl.h>
-#ifdef CBREAK
-# include <sgtty.h>
-# define SGTTY
-#else
-# ifdef TCGETA
-#  define TERMIO
-#  include <termio.h>
-# else
-#  define TERMIOS
-#  include <termios.h>
-# endif
-#endif
-#if defined(TERMIO) || defined(TERMIOS)
-# ifndef TAB3
-#  ifdef OXTABS
-#   define TAB3 OXTABS
-#  else
-#   define TAB3 0
-#  endif
-# endif
-#endif
-#include "screen.h"
-#include "boolean.h"
-
-extern char *myname;
-
-
-int  overstrike;
-int  screen_length;
-int  screen_width;
-char ch_erase;
-char ch_kill;
-char smart_terminal;
-char PC;
-char *tgetstr();
-char *tgoto();
-char termcap_buf[1024];
-char string_buffer[1024];
-char home[15];
-char lower_left[15];
-char *clear_line;
-char *clear_screen;
-char *clear_to_end;
-char *cursor_motion;
-char *start_standout;
-char *end_standout;
-char *terminal_init;
-char *terminal_end;
-
-#ifdef SGTTY
-static struct sgttyb old_settings;
-static struct sgttyb new_settings;
-#endif
-#ifdef TERMIO
-static struct termio old_settings;
-static struct termio new_settings;
-#endif
[…]

### 2017-02-22 downloads/top/top.c

@@ -1,1220 +1,0 @@
-char *copyright =
-    "Copyright (c) 1984 through 1996, William LeFebvre";
-
-/*
- *  Top users/processes display for Unix
- *  Version 3
- *
- *  This program may be freely redistributed,
- *  but this entire comment MUST remain intact.
- *
- *  Copyright (c) 1984, 1989, William LeFebvre, Rice University
- *  Copyright (c) 1989 - 1994, William LeFebvre, Northwestern University
- *  Copyright (c) 1994, 1995, William LeFebvre, Argonne National Laboratory
- *  Copyright (c) 1996, William LeFebvre, Group sys Consulting
- *
- * $FreeBSD: releng/11.0/contrib/top/top.c 300395 2016-05-22 04:17:00Z ngie $
- */
-
-/*
- *  See the file "Changes" for information on version-to-version changes.
- */
-
-/*
- *  This file contains "main" and other high-level routines.
- */
-
-/*
- * The following preprocessor variables, when defined, are used to
- * distinguish between different Unix implementations:
- *
- *	SIGHOLD  - use SVR4 sighold function when defined
- *	SIGRELSE - use SVR4 sigrelse function when defined
- *	FD_SET   - macros FD_SET and FD_ZERO are used when defined
- */
-
-#include "os.h"
-
-#include <sys/jail.h>
-#include <sys/time.h>
-
-#include <ctype.h>
-#include <errno.h>
-#include <jail.h>
-#include <setjmp.h>
-#include <signal.h>
-#include <unistd.h>
-
-/* includes specific to top */
-#include "commands.h"
-#include "display.h"		/* interface to display package */
-#include "screen.h"		/* interface to screen package */
-#include "top.h"
-#include "top.local.h"
-#include "boolean.h"
-#include "machine.h"
-#include "utils.h"
-#include "username.h"
-
-/* Size of the stdio buffer given to stdout */
-#define Buffersize	2048
-
-/* The buffer that stdio will use */
-char stdoutbuf[Buffersize];
-
-/* build Signal masks */
-#define Smask(s)	(1 << ((s) - 1))
-
-/* for getopt: */
-extern int  optind;
-extern char *optarg;
-
-/* imported from screen.c */
-extern int overstrike;
-
-static int fmt_flags = 0;
-int pcpu_stats = No;
-
-/* signal handling routines */
-sigret_t leave();
-sigret_t tstop();
[…]

### 2017-02-22 downloads/top/top.h

@@ -1,54 +1,0 @@
-/*
- * $FreeBSD: releng/11.0/contrib/top/top.h 300395 2016-05-22 04:17:00Z ngie $
- */
-/*
- *  Top - a top users display for Berkeley Unix
- *
- *  General (global) definitions
- */
-
-#ifndef TOP_H
-#define TOP_H
-
-/* Current major version number */
-#define VERSION		3
-
-/* Number of lines of header information on the standard screen */
-extern int Header_lines;	/* 7 */
-
-/* Maximum number of columns allowed for display */
-#define MAX_COLS	512
-
-/* Log base 2 of 1024 is 10 (2^10 == 1024) */
-#define LOG1024		10
-
-char *itoa();
-char *itoa7();
-
-char *version_string();
-
-/* Special atoi routine returns either a non-negative number or one of: */
-#define Infinity	-1
-#define Invalid		-2
-
-/* maximum number we can have */
-#define Largest		0x7fffffff
-
-/*
- * The entire display is based on these next numbers being defined as is.
- */
-
-#define NUM_AVERAGES    3
-
-enum displaymodes { DISP_CPU = 0, DISP_IO, DISP_MAX };
-
-/*
- * Format modifiers
- */
-#define FMT_SHOWARGS 0x00000001
-
-extern enum displaymodes displaymode;
-
-extern int pcpu_stats;
-
-#endif /* TOP_H */

### 2017-02-22 downloads/top/username.c

@@ -1,195 +1,0 @@
-/*
- *  Top users/processes display for Unix
- *  Version 3
- *
- *  This program may be freely redistributed,
- *  but this entire comment MUST remain intact.
- *
- *  Copyright (c) 1984, 1989, William LeFebvre, Rice University
- *  Copyright (c) 1989, 1990, 1992, William LeFebvre, Northwestern University
- *
- * $FreeBSD: releng/11.0/contrib/top/username.c 300395 2016-05-22 04:17:00Z ngie $
- */
-
-/*
- *  Username translation code for top.
- *
- *  These routines handle uid to username mapping.
- *  They use a hashing table scheme to reduce reading overhead.
- *  For the time being, these are very straightforward hashing routines.
- *  Maybe someday I'll put in something better.  But with the advent of
- *  "random access" password files, it might not be worth the effort.
- *
- *  Changes to these have been provided by John Gilmore (gnu@toad.com).
- *
- *  The hash has been simplified in this release, to avoid the
- *  table overflow problems of previous releases.  If the value
- *  at the initial hash location is not right, it is replaced
- *  by the right value.  Collisions will cause us to call getpw*
- *  but hey, this is a cache, not the Library of Congress.
- *  This makes the table size independent of the passwd file size.
- */
-
-#include <sys/param.h>
-#include <sys/types.h>
-
-#include <pwd.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-
-#include "top.local.h"
-#include "utils.h"
-#include "username.h"
-
-struct hash_el {
-    int  uid;
-    char name[MAXLOGNAME];
-};
-
-#define    is_empty_hash(x)	(hash_table[x].name[0] == 0)
-
-/* simple minded hashing function */
-/* Uid "nobody" is -2 results in hashit(-2) = -2 which is out of bounds for
-   the hash_table.  Applied abs() function to fix. 2/16/96 tpugh
-*/
-#define    hashit(i)	(abs(i) % Table_size)
-
-/* K&R requires that statically declared tables be initialized to zero. */
-/* We depend on that for hash_table and YOUR compiler had BETTER do it! */
-struct hash_el hash_table[Table_size];
-
-
-void
-init_hash()
-
-{
-    /*
-     *  There used to be some steps we had to take to initialize things.
[…]

### 2017-02-22 downloads/top/username.h

@@ -1,23 +1,0 @@
-/*
- *  Top users/processes display for Unix
- *  Version 3
- *
- *  This program may be freely redistributed,
- *  but this entire comment MUST remain intact.
- *
- *  Copyright (c) 1984, 1989, William LeFebvre, Rice University
- *  Copyright (c) 1989, 1990, 1992, William LeFebvre, Northwestern University
- *  Copyright (c) 2016, Randy Westlund
- *
- * $FreeBSD: releng/11.0/contrib/top/username.h 300395 2016-05-22 04:17:00Z ngie $
- */
-#ifndef USERNAME_H
-#define USERNAME_H
-
-int	 enter_user(int uid, char *name, int wecare);
-int	 get_user(int uid);
-void	 init_hash(void);
-char 	*username(int uid);
-int 	 userid(char *username);
-
-#endif /* USERNAME_H */

### 2017-02-22 downloads/top/utils.c

@@ -1,508 +1,0 @@
-/*
- *  Top users/processes display for Unix
- *  Version 3
- *
- *  This program may be freely redistributed,
- *  but this entire comment MUST remain intact.
- *
- *  Copyright (c) 1984, 1989, William LeFebvre, Rice University
- *  Copyright (c) 1989, 1990, 1992, William LeFebvre, Northwestern University
- *
- * $FreeBSD: releng/11.0/contrib/top/utils.c 278560 2015-02-11 07:44:53Z jmg $
- */
-
-/*
- *  This file contains various handy utilities used by top.
- */
-
-#include "top.h"
-#include "os.h"
-
-int atoiwi(str)
-
-char *str;
-
-{
-    register int len;
-
-    len = strlen(str);
-    if (len != 0)
-    {
-	if (strncmp(str, "infinity", len) == 0 ||
-	    strncmp(str, "all",      len) == 0 ||
-	    strncmp(str, "maximum",  len) == 0)
-	{
-	    return(Infinity);
-	}
-	else if (str[0] == '-')
-	{
-	    return(Invalid);
-	}
-	else
-	{
-	    return(atoi(str));
-	}
-    }
-    return(0);
-}
-
-/*
- *  itoa - convert integer (decimal) to ascii string for positive numbers
- *  	   only (we don't bother with negative numbers since we know we
- *	   don't use them).
- */
-
-				/*
-				 * How do we know that 16 will suffice?
-				 * Because the biggest number that we will
-				 * ever convert will be 2^32-1, which is 10
-				 * digits.
-				 */
-_Static_assert(sizeof(int) <= 4, "buffer too small for this sized int");
-
-char *itoa(val)
-
-register int val;
-
-{
-    register char *ptr;
-    static char buffer[16];	/* result is built here */
-    				/* 16 is sufficient since the largest number
-				   we will ever convert will be 2^32-1,
-				   which is 10 digits. */
-
-    ptr = buffer + sizeof(buffer);
-    *--ptr = '\0';
-    if (val == 0)
-    {
-	*--ptr = '0';
-    }
-    else while (val != 0)
-    {
-	*--ptr = (val % 10) + '0';
-	val /= 10;
-    }
-    return(ptr);
-}
-
-/*
- *  itoa7(val) - like itoa, except the number is right justified in a 7
- *	character field.  This code is a duplication of itoa instead of
- *	a front end to a more general routine for efficiency.
- */
-
-char *itoa7(val)
-
[…]

### 2017-02-22 downloads/w/extern.h

@@ -1,38 +1,0 @@
-/*-
- * Copyright (c) 1993
- *	The Regents of the University of California.  All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 4. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE REGENTS AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE REGENTS OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- *
- *	@(#)extern.h	8.1 (Berkeley) 6/6/93
- * $FreeBSD: releng/11.0/usr.bin/w/extern.h 243183 2012-11-17 16:47:05Z ed $
- */
-
-extern	int use_ampm;
-
-struct kinfo_proc;
-int	pr_attime(time_t *, time_t *);
-int	pr_idle(time_t);
-int	proc_compare(struct kinfo_proc *, struct kinfo_proc *);

### 2017-02-22 downloads/w/pr_time.c

@@ -1,130 +1,0 @@
-/*-
- * Copyright (c) 1990, 1993, 1994
- *	The Regents of the University of California.  All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 4. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE REGENTS AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE REGENTS OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <sys/cdefs.h>
-
-__FBSDID("$FreeBSD: releng/11.0/usr.bin/w/pr_time.c 274151 2014-11-05 23:54:33Z marcel $");
-
-#ifndef lint
[…]

### 2017-02-22 downloads/w/proc_compare.c

@@ -1,121 +1,0 @@
-/*-
- * Copyright (c) 1990, 1993
- *	The Regents of the University of California.  All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 4. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE REGENTS AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE REGENTS OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#if 0
-#ifndef lint
-static char sccsid[] = "@(#)proc_compare.c	8.2 (Berkeley) 9/23/93";
-#endif /* not lint */
-#endif
-
-#include <sys/cdefs.h>
[…]

### 2017-02-22 downloads/w/w.c

@@ -1,559 +1,0 @@
-/*-
- * Copyright (c) 1980, 1991, 1993, 1994
- *	The Regents of the University of California.  All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions and the following disclaimer.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 4. Neither the name of the University nor the names of its contributors
- *    may be used to endorse or promote products derived from this software
- *    without specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE REGENTS AND CONTRIBUTORS ``AS IS'' AND
- * ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED.  IN NO EVENT SHALL THE REGENTS OR CONTRIBUTORS BE LIABLE
- * FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
- * DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
- * OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
- * HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
- * LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
- * OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
- * SUCH DAMAGE.
- */
-
-#include <sys/cdefs.h>
-
[…]

### 2017-02-22 loadgen/loadgen.c

@@ -1,86 +1,0 @@
-/* loadgen.c 
- * author: Austin Voecks
- *
- * Program that simluates CPU load by doing some math and then waiting for an
- * empirically determined number of nanoseconds. 
- */
-
-#include <math.h>
-#include <time.h>
-#include <stdio.h>
-#include <stdlib.h>
-
-int main(int argc, char **argv){
-
-  int loads[100] = {0};
-  int counter, load, i;
-  double elapsed, duration;
-  float x, y, z;
-  time_t start, end;
-  struct timespec tim, tim2;
-
-  /* supported loads */
-  loads[15] = 30000;
-  loads[25] = 20000;
-  loads[50] = 9000;
-  loads[75] = 8100;
-  loads[95] = 7800;
-
-  if (argc != 3) {
-    printf("usage: %s [cpu load %%] [duration]\n", argv[0]);
-    return -1;
-  }
-
-  /* get and verify parameters */
-  load = atoi(argv[1]);
-  duration = (double)atoi(argv[2]);
-
-  if(load < 0 || load > 100 || duration < 1 ) {
-    printf("usage: %s [cpu load %%] [duration]\n", argv[0]);
-    return -1;
-  }
-
-  if (loads[load] == 0) {
-    printf("supported percentages: ");
-    for (i = 0; i < 100; i++)
-      if (loads[i] != 0)
-        printf("%d, ", i);
-    printf("\n");
-    return -1;
-  }
-
-  /* pre loop setup */
-  counter = 0;
-  tim.tv_sec = 0;
-  tim.tv_nsec = loads[load];
-  time(&start);
-  printf("starting %d (%d) percent load for %d seconds\n", 
-      load, loads[load], (int)duration);
-
-  while (elapsed < duration) {
-
-    /* work */
-    if (x < 1 || y < 1 || z < 1)
-      x = y = z = 0xffffffff;
-
-    for (i = 0; i < 10; i++) {
-      x = sqrt(x); x = sqrt(x); x = sqrt(x);
-      y = sqrt(y); y = sqrt(y); y = sqrt(y);
-      z = sqrt(z); z = sqrt(z); z = sqrt(z);
-    }
-
-    /* check times less frequently so we focus on the math */
-    if (counter == 50) {
-      time(&end);
-      elapsed = difftime(end, start);
-      counter = 0;
-    }
-    else {
-      counter += 1;
-    }
-
-    if(nanosleep(&tim , &tim2) < 0 )
-      return -1;
-  }
-  return 0;
-}

### 2017-02-22 scripts/cpu_usage.sh

@@ -2,8 +2,8 @@
 
 finish() {
   echo
-  echo $results
-  echo $results | awk '{s+=$1}END{print "average:",s/NR}' RS=" "
+  echo "$results"
+  awk '{s+=$1}END{print "average:",s/NR}' RS=" " <<< "$results"
   exit
 }
 
@@ -11,8 +11,8 @@ trap finish INT
 results=""
 
 while true; do
-  pid="$(pgrep $1 | head -n 1)"
-  if test "$pid" != ""; then
+  pid="$(pgrep "$1" | head -n 1)"
+  if ! [[ -z "$pid" ]]; then
     results="$results $(ps -p "$pid" -o '%cpu' | tail -n 1 | grep -o '[0-9\.]*')"
   fi
   sleep 0.5

### 2017-02-22 scripts/make_rules.sh

@@ -11,7 +11,7 @@ fi
 base=5000
 
 for i in $(seq "$1"); do
-  num=$(( $i + $base ))
+  num=$(( i + base ))
   ipfw -q add "$num" deny ip from 127.255.0."$i" to any
 done
 

### 2017-02-22 paper/main.tex
Update paper, final draft

@@
 \author{Philip Nelson and Austin Voecks% <-this % stops a space
+Western Washington University
 \thanks{Supported in part by Dell EMC Isilon through grant and hardware}}
@@
 
-% TODO rewrite
 \section{Introduction}
 
-  IPFWD is a daemon for FreeBSD that updates an early rule in IPFW that has a
-  chance to accept any packet. The probability of acceptance is updated over
-  time and dependent on the current system load. In high performance systems
-  this results in an increase in network throughput.
+  Firewalls shield networks and hosts from malicious traffic by blocking
+  network packets that do not match any of the rules defined in a security
+  policy. 
+  In high performance and network throughput environments, extensive packet
+  processing by firewalls can limit the system's ability to utilize the full
+  capacity of network links. This problem is typically addressed by restricting
+  the number and complexity of firewall rules, disabling the firewall entirely,
+  or sacrificing the additional throughput so all the packet processing can be
+  completed.
+
+%  The default action for white-list firewalls is to block all traffic,
+%  additional rules describe exceptions. Blacklist firewalls are the opposite,
+%  by default they accept all traffic and additional rules describe packet types
+%  that are not allowed. Since predicting and blacklisting all possible attacks
+%  is not possible in practice, white-list firewall rule sets are the accepted
+%  standard and the focus of this paper. Without a security policy,
+%  administrators cannot know what to allow and disallow when constructing
+%  firewall rules.
+
+    % Packet sampling
+    Other papers \cite{exploitingpacketsampling, analysisnetflow,
+    monitoringpacketsampling} on packet sampling have been focused on detecting
+    and classifying threats, not performance enhancement. However, their
+    approaches to packet sampling and the implications addressed provided the
+    inspiration for the work done here. The BSD community has produced several in depth investigations into network
+    and firewall performance \cite{ipfwvspf,optimizingfreebsd}. Their work
+    informed the tests run here.
+
+\section{IPFW \& IPFWD}
+
+  IPFWD is a daemon for FreeBSD that updates an early rule in IPFW (the IP
+  Firewall) that has a chance to accept any packet. The probability of early
+  acceptance is updated over time and dependent on the current system load. In
+  high performance systems this can result in an increase in network
+  throughput.
 
@@
 
-\section{IPFW}
-
-  % Firewall basics
-  \subsection{Firewall Basics}
-
-    Firewalls restrict the types of incoming and outgoing network traffic
-    according to a security policy. Without a security policy, administrators
-    cannot know what to allow and disallow. Typical firewalls enforce the
-    policy by blocking packets that do not match the rules. The default action
-    for white-list firewalls is to block all traffic, additional rules describe
-    exceptions. Blacklist firewalls are the opposite in that by default they
-    accept all traffic and additional rules describe packet types that are not
-    allowed. Since predicting and blacklisting all possible attacks is
-    infeasible, white-list firewall rule sets are the accepted standard and the
-    focus of this paper.
-
   % Why was this firewall chosen?
@@
 
-    IPFW was chosen for it's close integration with the FreeBSD operating
-    system, in kernel packet filter, and general performance. Performance is
-    dependent on system parameters and circumstances, but IPFW outperforms PF
-    in stateful packet filtering \cite{ipfwvspf}. Stateful rules are more
-    powerful and flexible than stateless rules, were no session information is
-    maintained \cite{networksecurity}.  Additionally, the capabilities of IPFW
-    extend beyond packet filtering into source based routing, traffic shaping
-    and more.
+    IPFW was chosen for its close integration with the FreeBSD operating
+    system, kernel packet filter, and performance. Performance is dependent on
+    system parameters and circumstances, but IPFW outperforms the other main
+    firewall option for FreeBSD, PF, in stateful packet filtering
+    \cite{ipfwvspf}. Stateful rules are more powerful and flexible than
+    stateless rules, where no session information is maintained
+    \cite{networksecurity}.  Additionally, the capabilities of IPFW extend
+    beyond packet filtering into source based routing, traffic shaping and
+    more. It's for these reasons that IFPW was chosen over PF.
 
@@
 
-    This paper is not meant as a comparison between IPFW and PF; the
-    functionality provided by IPFWD could be ported to PF and serves as a proof
-    of concept for firewalls in general.
-
 \section{IPFWD}
@@
   It does this by adding and maintaining an early rule in the IPFW rule set of
-  the following form: \verb|prob 0.000 allow ip from any to any|.
+  the following form: 
+  \begin{center}
+      \verb|prob 0.000 allow ip from any to any|
+  \end{center}
+  which allows any IP packet, incoming or outgoing with a chance equal to the
+  probability given.
 
@@
   large amounts of rejected traffic; normally each rejected packet has all
-  rules applied to it while IPFW looks for a match.  When no matches are
+  rules applied to it in order to find a match. When no matches are
   encountered, the default rule is employed and the packet is rejected.
 
-  In a typical data center backend network the vast majority of traffic is
-  valid and going to be accepted by the firewall's rule set. However, these
-  rule sets are often complex and take increasing time to match packets
-  depending on the number of rules. For a rule set of length $n$, we can expect
-  to a packet to encounter $\frac{n}{2}$ failed matches before matching the
-  correct rule and being accepted.
+  In a typical data center internal network, the vast majority of traffic is
+  valid and going to be accepted by the firewalls rule set. However, these rule
+  sets are often complex and take increasing time to match packets depending on
+  the number of rules. For a rule set of length $n$, we can expect a packet to
+  encounter on average $\frac{n}{2}$ failed matches before matching the correct
+  rule and being accepted.
 
@@
 
-  where $k$ is number of rules before the early acceptance rule and
-  $\mathsf{P}(\textit{EA})$ is the probability that the early acceptance rule
-  is matched.
+  where $k$, a small constant compared to $n$, is number of rules before the
+  early acceptance rule and $\mathsf{P}(\textit{EA})$ is the probability that
+  the early acceptance rule is matched.
 
-  Removing the constant $k$, IPFWD reduces the number of failed matches before
+  Simplifying in order notation, IPFWD reduces the number of failed matches before
   a successful match by a factor of
@@
   creates a chance that packets that would have normally been rejected will get
-  through the firewall. We'll refer to these types of packets as invalid for
-  the purpose of discussion.
+  through the firewall.
 
-  In order to provide visibility into this process, IPFWD extends the IPFW
-  system logs to include information about the early acceptance probability.
-  From this information, administrators can extrapolate the probability that
-  additional invalid packets made it through the firewall for each that is
-  detected. This is discussed further in the \textit{Security Implications}
-  section.
+  In order to provide visibility into this process, IPFWD writes its own logs
+  to include information about the early acceptance probability. From this
+  information, administrators can extrapolate the probability that additional
+  invalid packets made it through the firewall for each that is detected. This
+  is discussed further in \ref{securityimplications}.
 
@@
 
-        The gains provided by IPFWD are only apparent in systems were network
+        The gains provided by IPFWD are only apparent in systems where network
         throughput is matched by or out performs CPU speed. Systems that can
@@
         reducing processing time. Given modern CPU clock speeds, systems with
-        network cards slower than 1 Gb/sec will likely not benefit either.
-
-      \item \textbf{Resource Conscious}
-
-        In line with high performance systems, systems with strict resource
-        restrictions can benefit by the reduced network related CPU usage.
+        network cards slower than 1 Gb/sec will likely not benefit from IPFWD.
 
@@
 
-\section{Security Implications}
+\section{Security Implications} \label{securityimplications}
 
   % Talk about security policy enforcement
-  Security policies define what is and is not allowed in an organization . At
+  Security policies define what is and is not allowed in an organization. At
   the network and host level, this is implemented in part by a firewall rule
@@
 
-  Tests were run on three platforms and five different network cards, resulting
-  in five different systems. The most important parameters to our tests are
-  CPU, NIC, and ENV. Descriptions of each system are given in the tables
-  below:\\
+  We benchmarked the performance of IPFW in relation to CPU usage and network
+  throughput. The goal was to determine if IPFWD reduces the average time IPFW
+  takes to process packets. Tests were run on three platforms and four
+  different network cards, referred to as four different systems. The most
+  important parameters to these tests are CPU, NIC, and ENV. Descriptions of
+  each system are given in the tables below:\\
 
@@
     Multiple network performance tools were explored but the final results were
-    compiled using Netperf. Netperf was chosen because it was readily available
-    on all test platforms, provides built-in CPU utilization measurement, and a
-    wide range of test types \cite{netperf}.
+    compiled using Netperf. Netperf generates network traffic between client
+    and server instances and measures throughput, errors, and other network
+    performance metrics. Netperf was chosen because it was readily available
+    on all test platforms, provides built-in CPU utilization measurement, has a
+    wide range of test types \cite{netperf}, and is used elsewhere in the BSD
+    community for testing.
 
@@
     % Experimental method
-    All tests were conducted using the following steps:
-    \begin{enumerate}
-        \item Machine 1
-        \begin{itemize}
-          \item \texttt{ipfw}
-          \item \texttt{netserver}
-          \item \texttt{CPU monitor}
-        \end{itemize}
-        \item Machine 2
-        \begin{itemize}
-          \item \texttt{ipfw}
-          \item \texttt{ipfwd}
-          \item \texttt{netperf}
-          \item \texttt{CPU monitor}
-        \end{itemize}
-    \end{enumerate}
-
-    All tests were run with minimal IPFW rule sets in a range of 30 to 60 rules
-    depending on the environment and network card configuration. Larger
-    production rule sets would benefit even further from IPFWD.
+    All tests were conducted using the following setup:
+    \begin{center}
+    \begin{tabular}{ | L{10em} | L{10em} | } 
+      \hline
+      Machine 1 & Machine 2 \\
+      \hline
+      \hline
+      \texttt{ipfw}        & \texttt{ipfw} \\
+      \texttt{netserver}   & \texttt{ipfwd} \\
+      \texttt{CPU monitor} & \texttt{netperf} \\
+                           & \texttt{CPU monitor} \\
+      \hline
+    \end{tabular} \\
+    \end{center}
 
@@
 
-    CPU utilization measurement is the most difficult parameter to measure in
+    CPU utilization measurement was the most difficult parameter to measure in
     these tests. There are many factors that can account for variability
     between tests and jitter in the results. The Netperf manual describes these
-    challenges and the variety of techniques used on different systems.
-    \cite{netperf}
+    challenges and the variety of techniques used on different systems to
+    measure CPU usage \cite{netperf}.
 
-    Netperf was run in single threaded mode, which explains why throughput
-    wasn't closer to the maximum throughput for the NIC on each system. The
+    Netperf was run in single threaded mode, which explains why throughput was
+    not closer to the maximum throughput for the NIC on each system. The
     FreeBSD network stack allows multithreaded packet processing by utilizing
-    multiple queues \cite{freebsdhandbook}. Since we're indirectly testing the
-    performance of IPFW through Netperf, there are two different CPU
+    multiple queues meaning the kernel can spread the work done by IPFW over
+    multiple cores \cite{freebsdhandbook}. Since we were indirectly testing the
+    performance of IPFW through Netperf, there were two different CPU
     measurements to account for: kernel time and user time. Since IPFW runs in
-    the kernel, it's CPU time is classified as kernel time.
+    the kernel, its CPU time is classified as kernel time.
 
-    Netperf was able to measure CPU utilization on the Virtualized and WWU
+    Netperf was able to measure CPU utilization on the Virtualised and WWU
     systems, but didn't work on the Isilon systems. Instead, \textit{ps} was
-    used. These tests are comparable since both measured user CPU utilization,
-    not kernel CPU utilization.
-
-    In systems where the throughput of the NIC exceeds what the CPU can supply,
-    CPU utilization is high and throughput is lower than the NIC maximum (eg:
-    Virtualised system). Conversely, when CPU availablility exceeds NIC
-    throughput, throughput is high and CPU utilization is low (eg: WWU 1G
-    system).
+    used. \texttt{ps} is a BSD utility that displays information about
+    processes, including CPU utilization. We used \texttt{ps} to poll Netperf's
+    CPU usage and averaged the output to get the results presented here. These
+    tests are comparable since both measured user CPU utilization, not kernel
+    CPU utilization. Thus, IPFWs CPU usage was measured indirectly through the
+    CPU usage of netperf.
 
@@
 
-    These graphs show the interaction between CPU utilization and network
-    throughput for each system with respect to IPFWD's early acceptance
-    probability. The dashed lines are CPU utilization, solid is network
-    throughput in Mb/second. Higher acceptance probability means a greater
-    chance that a given packet is immediately accepted without further
-    processing.
+    The graphs in figure \ref{fig:isiloncpu40}, \ref{fig:isiloncpu10},
+    \ref{fig:wwucpu1}, and \ref{fig:virtualcpu10} show the interaction between
+    CPU utilization and network throughput for each system with respect to
+    IPFWDs early acceptance probability. The dashed lines are CPU utilization;
+    solid is network throughput in Mb/second. A higher acceptance probability
+    means a greater chance that a given packet was immediately accepted without
+    further processing.
 
@@
     acceptance probability increases, so does throughput. The CPU usage shows
-    that this throughput is gained without additional CPU utilization, meaning
-    the system is spending fewer cycles to process each packet on average.
+    that this additional throughput was gained without additional CPU
+    utilization, meaning the system was spending fewer cycles to process each
+    packet on average.
 
-    The lack of any trend in the WWU 1G graph is expected and addressed in the
-    \textit{Discussion} section.
+    The lack of any trend in the WWU 1G graph \ref{fig:wwucpu1} was expected
+    and will be addressed in \ref{discussion}.
 
-    \begin{center}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{cpu_isilon40}
+      \caption{Throughput \& CPU for Isilon 40G system}
+      \label{fig:isiloncpu40}
+    \end{figure}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{cpu_isilon10}
+      \caption{Throughput \& CPU for Isilon 10G system}
+      \label{fig:isiloncpu10}
+    \end{figure}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{cpu_wwu1}
+      \caption{Throughput \& CPU for WWU 1G system}
+      \label{fig:wwucpu1}
+    \end{figure}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{cpu_virtual10}
-    \end{center}
+      \caption{Throughput \& CPU for Virtualised 10G system}
+      \label{fig:virtualcpu10}
+    \end{figure}
 
@@
 
-    These graphs show network the ratio of throughput to CPU utilization for
-    each system where $ratio(x) = throughput(x) / cpu(x)$. The data being
-    presented here is the same as the previous graphs, shown differently. The
-    upward trend over the x-axis shows that as the probability of early
-    acceptance increases, the system is spending less CPU time to process
-    packets on average.
+    The graphs in figures \ref{fig:isilontoc40}, \ref{fig:isilontoc10},
+    \ref{fig:wwutoc1}, and \ref{fig:virtualtoc10}  show network the ratio of
+    throughput to CPU utilization for each system where $ratio(x) =
+    throughput(x) / cpu(x)$. The data being presented here is the same as the
+    previous graphs, merely shown differently. The upward trend over the x-axis
+    shows that as the probability of early acceptance increases, the system is
+    spending less CPU time to process packets on average.
 
     The dashed line shows the line of best fit for the throughput to CPU
-    utilization ratio and shows more clearly the upward trend.
+    utilization ratio and shows the upward trend more clearly than the raw
+    data.
 
-    \begin{center}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{toc_isilon40}
+      \caption{Throughput/CPU for Isilon 40G system}
+      \label{fig:isilontoc40}
+    \end{figure}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{toc_isilon10}
+      \caption{Throughput/CPU for Isilon 10G system}
+      \label{fig:isilontoc10}
+    \end{figure}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{toc_wwu1}
+      \caption{Throughput/CPU for WWU 1G system}
+      \label{fig:wwutoc1}
+    \end{figure}
+    \begin{figure}[h]
       \includegraphics[width=0.45\textwidth]{toc_virtual10}
-    \end{center}
+      \caption{Throughput/CPU for Virtualised 10G system}
+      \label{fig:virtualtoc10}
+    \end{figure}
 
   % Difference in results between CPU and throughput bound systems
-  \subsection{Discussion}
+  \section{Discussion} \label{discussion}
 
@@
 
-\section{Related Work}
+    In summary, systems where the throughput of the NIC exceeds what the CPU
+    can supply CPU utilization is high and throughput is lower than the NIC
+    maximum (eg: the Virtualised system \ref{fig:virtualcpu10}). Conversely,
+    when CPU availability exceeds NIC throughput, throughput is high and CPU
+    utilization is low (eg: the WWU 1G system \ref{fig:wwucpu1}).
 
-  % Packet sampling
-  Other papers \cite{exploitingpacketsampling, analysisnetflow,
-  monitoringpacketsampling} on packet sampling have been focused on detecting
-  and classifying threats, not performance enhancement. However, their
-  approaches to packet sampling and the implications addressed provided the
-  idea for IPFWD.
-
-  % IPFW performance
-  The BSD community has produced several in depth investigations into network
-  and firewall performance \cite{ipfwvspf,optimizingfreebsd}. Their work
-  informed the tests run to measure IPFWD's performance benefits.
+    Additionally, all tests were run with minimal IPFW rule sets in a range of
+    30 to 60 rules depending on the environment and network card configuration.
+    Larger production rule sets would benefit even further from IPFWD.
 
@@
   % Firewall rule reordering so we can "fail as early as possible"
-  Ideally, a white-list firewall's rule set would be constructed so that the
-  most commonly applied matches are reached earlier rather than later. IPFWD
+  Ideally, a white-list firewalls rule set would be constructed so that the
+  most commonly applied matches are reached sooner rather than later. IPFWD
   accomplishes this task by artificial early stopping. Firewall rule reordering
-  is a could be another solution to this problem. Given logs of common network
+  could be another solution to this problem. Given logs of common network
   traffic for the system, statistical analysis could show which rules are
   matched more often than others. With this information, the rule set could be
-  reordered so that the most commonly applied rules are earliest.
-
-  Other work \cite{redundancydetection} has been done in removing redundant
-  rules to increase firewall performance. It's possible that a combination of
-  redundancy checking and priority reordering could provide further performance
-  gains.
+  reordered so that the most commonly applied rules are earliest. A combination
+  of redundancy checking and priority reordering could provide further
+  performance gains as well.
 
@@
   % see any invalid packets 
-  Machine learning could be applied to packet samples to more accurately
+  Further analysis could be done to packet samples in order to more accurately
   determine appropriate bounds on the early acceptance probability. Different
-  systems and circumstances may require more or less stringent packet checking.
+  systems and environments may require more or less stringent packet checking.
   Environments with very high confidence in network security might raise the
@@
 
+  % mutli threaded netperf tests
+  The tests conducted here were restricted to one core; future work could
+  explore the interaction of early acceptance on parallel firewall packet
+  processing. Running additional, concurrent netperf tests would show how the
+  system behaves when the links are fully saturated.
+
   % have detection of bad traffic reduce acceptance rate
@@
   The source code for IPFWD, additional test results, and development cycle
-  information may be found at \texttt{https://???}.
+  information may be found at \texttt{https://github.com/Gandalf-/ipfwd}.
 
-% use section* for acknowledgement
-%\section*{Acknowledgment}
-% TODO
+\section*{Acknowledgment}
+
+The authors would like to thank Dell EMC Isilon for providing hardware, access
+to clusters for testing, and funding for this research. The
+resources and the advice on FreeBSD network capabilities/network performance
+testing were invaluable. A portion of this work was done during a summer
+internship at Dell EMC Isilon.
 

### 2017-02-22 paper/main.tex
Revised draft

@@
   network packets that do not match any of the rules defined in a security
-  policy. 
-  In high performance and network throughput environments, extensive packet
-  processing by firewalls can limit the system's ability to utilize the full
-  capacity of network links. This problem is typically addressed by restricting
-  the number and complexity of firewall rules, disabling the firewall entirely,
-  or sacrificing the additional throughput so all the packet processing can be
-  completed.
+  policy. In high performance and network throughput environments, extensive
+  packet processing by firewalls can limit the system's ability to utilize the
+  full capacity of network links. This problem is typically addressed by
+  restricting the number and complexity of firewall rules, disabling the
+  firewall entirely, or sacrificing the additional throughput so all the packet
+  processing can be completed.
 
-%  The default action for white-list firewalls is to block all traffic,
-%  additional rules describe exceptions. Blacklist firewalls are the opposite,
-%  by default they accept all traffic and additional rules describe packet types
-%  that are not allowed. Since predicting and blacklisting all possible attacks
-%  is not possible in practice, white-list firewall rule sets are the accepted
-%  standard and the focus of this paper. Without a security policy,
-%  administrators cannot know what to allow and disallow when constructing
-%  firewall rules.
-
-    % Packet sampling
-    Other papers \cite{exploitingpacketsampling, analysisnetflow,
+  Our work uses a form of packet sampling to provide a trade-off between
+  complete application of a security policy and increased performance.  Other
+  papers \cite{exploitingpacketsampling, analysisnetflow,
   monitoringpacketsampling} on packet sampling have been focused on detecting
@@
   approaches to packet sampling and the implications addressed provided the
-    inspiration for the work done here. The BSD community has produced several in depth investigations into network
-    and firewall performance \cite{ipfwvspf,optimizingfreebsd}. Their work
-    informed the tests run here.
+  inspiration for the work done here. The BSD community has produced several in
+  depth investigations into network and firewall performance
+  \cite{ipfwvspf,optimizingfreebsd}. Their work informed the tests run here.
 
@@
   high performance systems this can result in an increase in network
-  throughput.
+  throughput without additional CPU load.
 
